---
title: "BBS_neonics_landuse"
author: "Olivia Burge"
date: "21 January 2015"
output: html_document
---

````{r, warning = FALSE, message = FALSE, results='hide'}
require(ggthemes)
require(wesanderson)
require(ggplot2)
require(vegan)
require(dplyr)
require(data.table)
require(tidyr)
````

# Reading in the new bird data
From Tyson's csv. Creates a column for the abundance of birds by route (data is by stop).
````{r}
setwd("/Users/Liv/Documents/NCEAS_GIT/NCEAS-RENCI_2014/BBS_data/")
new_bbs <- read.csv("fifty7_withRTENO.csv")# 
new_bbs$sum_route_abundance <- rowSums(new_bbs[grep("^Stop[0-9]+", names(new_bbs))])
head(new_bbs)
names(new_bbs)
new_bbs <- new_bbs[ , c(1:7, 58, 59)] #selects everything except the individual stop abundances
length(unique(new_bbs$Route))
head(new_bbs)
````

# Bird trait data
Read in and pare down the bird trait data to the scientific, common names for species and their habitat, nesting, diest, behaviour and conservation status, and 4&6 letter codes.

````{r}
bird_traits <- read.csv("OH_BBS_1966-2013_traits.csv")
bird_traits <- bird_traits[ , c(1,2,3,4,5,6,7,8,9)]
head(bird_traits[2])
names(bird_traits) <- toupper(names(bird_traits))
names(bird_traits) <- gsub("[.]", "_", names(bird_traits))  # using the [] to get the period to be used literally. Or can use "fixed = TRUE"
names(bird_traits)
bird_traits$SCIENTIFIC_NAME <- toupper(bird_traits$SCIENTIFIC_NAME)
head(bird_traits)
````

# Bird code AOU data
BBS data comes with a different kind of AOU code. This chunk tidies up the AOU code dataframe and makes the common and scienitific names compatible with the BBS data. Need to use the common names, scientific names are slightly off in some cases.

```{r}
AOU_codes <- read.csv("raw_data/AOU_codes.csv") ## a cleaned up (ie header removed, split into columns - no substantive changes to any species names or column headers) version of: ftp://ftpext.usgs.gov/pub/er/md/laurel/BBS/DataFiles/SpeciesList.txt

head(AOU_codes)
AOU_codes$common_name <- toupper(AOU_codes$English_Common_Name)
names(AOU_codes)
AOU_codes <- AOU_codes[ , c(2, 10, 6:9)]
names(AOU_codes) <- toupper(names(AOU_codes))
head(AOU_codes)
AOU_codes$SCIENTIFIC_NAME <- paste(AOU_codes$GENUS, AOU_codes$SPECIES, sep = " ")
AOU_codes$SCIENTIFIC_NAME <- toupper(AOU_codes$SCIENTIFIC_NAME)
#testing that all our traits at leat are in the new AOU codes:
bird_traits$SCIENTIFIC_NAME %in% AOU_codes$SCIENTIFIC_NAME # = not all true.
AOU_codes$SCIENTIFIC_NAME %in% bird_traits$SCIENTIFIC_NAME # many many falses
bird_traits$COMMON_NAME <- toupper(bird_traits$COMMON_NAME)
head(bird_traits)
bird_traits$COMMON_NAME %in% AOU_codes$COMMON_NAME # = all true!
AOU_codes$English_Common_Name %in% bird_traits$COMMON_NAME # many many falses

unique(new_bbs$AOU) %in% unique(AOU_codes$AOU) # good.

# merge the AOU codes into the ohio data
```

# Merging AOU codes and traits

Putting the trait and AOU codes together - 
```{r}
head(AOU_codes)
head(new_bbs)

traits_aou <- inner_join(bird_traits, AOU_codes)

BBS_traits <- inner_join(new_bbs, traits_aou)
names(BBS_traits)
````

# Neonics data by route
Based on Tyson's GIS analysis.
````{r}
neonics_buffers <- read.csv("../Pesticides/pest_buff_overtime.csv")
neonics_buffers$NEONIC <- ifelse(neonics_buffers$COMPOUND %in% c("ACETAMIPRID", "CLOTHIANIDIN", "FIPRONIL", "IMIDACLOPRID", "THIACLOPRID", "THIAMETHOXAM"), "neonic", "pesticide")
head(neonics_buffers)
tail(neonics_buffers)

neonics_buffers_short <- neonics_buffers %>%
  group_by(YEAR, RTENO, buffer, NEONIC) %>%
  summarise(SUM_high_kg = sum(high_kg_buff))
head(neonics_buffers_short)
#neonics_buffers[is.na(neonics_buffers)]
# quite a few NAs in hte low kg buff col
````

# Merge BBS_traits and neonics
```{r}
sort(unique(neonics_buffers$YEAR))
names(BBS_traits) <- toupper(names(BBS_traits))
sort(unique(BBS_traits$YEAR))
````
Therefore, we need to select the period 1995 - 2011 for merging.

```{r}
BBS_neonics <- inner_join(BBS_traits[ BBS_traits$YEAR < 2012, ], neonics_buffers_short[ neonics_buffers$YEAR > 1994, ]) # joins by RTENO and YEAR which works. 
````

# Landuse data by route
Based on Tyson's GIS analysis. Notes for the csv: `reclass` is Anderson Level I land cover: (0=nodata, 1=water, 2=developed, 3=barren, 4=forest, 5=grassland/shrub, 6=agriculture, 7=wetlands).
`total` is the number of pixels (multiply by 0.009 to get square km).
````{r}
landuse_buffers <- read.csv("../Landcover/LC_buffers_overtime.csv")
head(landuse_buffers)

reclass_table <- data.frame("reclass" = 0:7, "land_cover" = c("nodata", "water", "developed", "barren", "forest", "grassland_shrub", "agriculture", "wetlands"))

landuse_buffers <- merge(landuse_buffers, reclass_table, by = "reclass")
class(landuse_buffers$land_cover)

head(landuse_buffers)
tail(landuse_buffers)

````
````{r}
names(landuse_buffers)

# dropping this second one as it doesn't have the old cols

# landuse_buffers2 <- landuse_buffers[c(2, 3, 4, 5, 7, 8)]
# names(landuse_buffers2)
# class(landuse_buffers2$total_pix)
# str(landuse_buffers2)
# head(landuse_buffers2)

landuse_buffers_wide <- spread(landuse_buffers[, c(2:5, 7)], land_cover, total_pix)

tail(subset(landuse_buffers_wide,YEAR == 2012), 50)
# we'll need to exclude the 2012 ata from here
names(landuse_buffers_wide)
````
# Merging BBS_neonics with landuse
```{r}
sort(unique(landuse_buffers$YEAR)) # 1992 - 2011



BBS_ln <- inner_join(BBS_neonics, landuse_buffers_wide[ landuse_buffers_wide$YEAR > 1994, ])
# joins by: c("YEAR", "RTENO", "BBS_route", "buffer")

tail(BBS_ln)
sort(unique(BBS_ln$YEAR))
```

# Maybe a baby model?

```{r}
names(BBS_ln)
unique(BBS_ln$buffer)


# 
# BBS_ln$agriculture_scaled <- as.numeric(scale(BBS_ln$agriculture))
# BBS_ln$high_kg_buff_scaled <- as.numeric(scale(BBS_ln$high_kg_buff))
head(BBS_ln)
levels(BBS_ln$COMPOUND)
BBS_ln_scale <- BBS_ln %>%
  group_by(buffer) %>%
  mutate(ag_scaled = as.numeric(scale(agriculture)), 
         forest_scaled = as.numeric(scale(forest)))

names(BBS_ln_scale)
names(BBS_ln_scale) <- toupper(names(BBS_ln_scale))
BBS_ln_scale <- BBS_ln_scale %>%
  ungroup()%>%
  group_by(NEONIC, BUFFER) %>%
  mutate(high_kg_buff_scaled = as.numeric(scale(SUM_HIGH_KG)))
names(BBS_ln_scale) <- toupper(names(BBS_ln_scale))
```

```{r}
levels(BBS_ln_scale$DIET)
BBS_ln_scale$DIET_VULN <- ifelse(BBS_ln$DIET %in% c("insects", "seeds", 'nectar', "birds"), "diet_vuln", "diet_invuln")
BBS_ln_scale$DIET_VULN <- factor(BBS_ln_scale$DIET_VULN)

BBS_ln_scale$DIET_INSECTS <- ifelse(BBS_ln_scale$DIET == "insects", "diet_insects", "diet_not_insects")
BBS_ln_scale$DIET_INSECTS <- factor(BBS_ln_scale$DIET_INSECTS)

names(BBS_ln_scale)
BBS_ln_scale <- data.frame(BBS_ln_scale)
head(BBS_ln_scale)

BBS_ln_scale2 <- BBS_ln_scale %>%
  ungroup() %>%
  group_by(YEAR, RTENO, BUFFER, AOU, 
          HABITAT, DIET, NESTING, BEHAVIOR, COMMON_NAME,
          CONSERVATION, AGRICULTURE, BARREN, 
          FOREST, DEVELOPED, GRASSLAND_SHRUB, WATER,
          WETLANDS, AG_SCALED, FOREST_SCALED,
          DIET_VULN, DIET_INSECTS, NEONIC) %>%
  dplyr::summarise(sum_abundance = sum(SUM_ROUTE_ABUNDANCE), 
                   sum_neonics = sum(SUM_HIGH_KG))
dim(BBS_ln_scale)
dim(BBS_ln_scale2)
BBS_ln_scale2[!complete.cases(BBS_ln_scale2),]
  
names(BBS_ln_scale2)
BBS_ln_scale2 <- BBS_ln_scale2 %>% 
  ungroup() %>%
  group_by(NEONIC, BUFFER) %>%
  mutate(scale_neonics = as.numeric(scale(sum_neonics))) %>%
  ungroup() %>%
  group_by(BUFFER) %>%
  mutate(year_scaled = as.numeric(scale(YEAR)))

##### next model -scaled and neonics 
names(BBS_ln_scale2)
unique(BBS_ln_scale2$BUFFER)
range(BBS_ln_scale2$RTENO)
BBS_ln_scale2$ROUTE_FAC<- as.factor(BBS_ln_scale2$RTENO)
BBS_ln_scale3 <- BBS_ln_scale2[BBS_ln_scale2$BUFFER == 200 , ]
BBS_ln_scale3$obs <- 1:nrow(BBS_ln_scale3)

unique(BBS_ln_scale3$NEONIC)
BBS_ln_scale3$NEONIC <- factor(BBS_ln_scale3$NEONIC,
                                  levels = c("pesticide", "neonic"))  # changes the levels
BBS_ln_scale3$NEONIC <- relevel(BBS_ln_scale3$NEONIC, ref = "pesticide") #changes the ereference level
names(BBS_ln_scale3)

### up to here
#save.image("ohhai.Rdata")
#load("ohhai.Rdata")

```

```{r multivariate}
names(BBS_ln)
sort(unique(BBS_ln$YEAR))

length(unique(BBS_ln$RTENO))
length(unique(BBS_ln[BBS_ln$YEAR == 1998,"RTENO"]))

routes <- BBS_ln %>% group_by(YEAR) %>% summarise(num_routes = length(unique((RTENO))))

routes_plot <- ggplot(routes, aes(x = YEAR, y = num_routes)) +
  geom_line() +theme_bw() + labs(y = "Number of UNIQUE routes counted") +
  geom_vline(aes(xintercept = 1998), colour = "darkblue",
             linetype = "dashed") +
  annotate("text", x = 1998.3, y = 20, 
           hjust = 0,
           label = "1998 = suggested date from \n which to take bird obs",
           colour = "darkblue") +
  ggtitle("BBS data: number of routes counted by year")
ggsave(plot = routes_plot, "unique routes by year.pdf",
       width = 210, height = 148.5, units = "mm")

### multivariate landuse #####

multi_env <- BBS_ln  %>% filter(YEAR %in% c(1998, 2011)) %>% 
  group_by(YEAR, buffer, RTENO, NEONIC) %>% 
  slice(1)  %>% 
  ungroup() %>%
  select(RTENO, buffer, NEONIC, SUM_high_kg, forest, agriculture, developed, water, barren, grassland_shrub, wetlands, YEAR) %>%
  spread(NEONIC, SUM_high_kg)

multi_env1000_11$RTENO %in% multi_env1000_98$RTENO


multi_env1000_98 <- multi_env %>%
  filter(YEAR ==1998)

multi_env1000_11 <- multi_env %>%
  filter(YEAR ==2011)

#buffers <- c(200, 400, 1000, 2000, 5000, 10000)
multi_env$buffer2 <- as.factor(multi_env$buffer)
multi_env <- data.frame(multi_env)

lists98 <- split(multi_env[multi_env$YEAR == 1998,], 
                 f = multi_env[multi_env$YEAR == 1998,"buffer"])
str(lists98)
lists11 <-  split(multi_env[multi_env$YEAR == 2011,], 
                 f = multi_env[multi_env$YEAR == 2011,"buffer"])

multi_env1000_98_pro <- multi_env1000_98[multi_env1000_98$RTENO %in% multi_env1000_11$RTENO, ] %>% arrange(RTENO)
multi_env1000_11_pro <- multi_env1000_11[multi_env1000_11$RTENO %in% multi_env1000_98_pro$RTENO , ] %>% arrange(RTENO)


save(lists98, file = "lists98.Rdata")
for (dt in seq(along = lists98)){
  metas98 = list()
  meta_n = metaMDS(lists98[[dt]][3:9])
  metas98[[dt]] <- meta_n
  names(metas98)[[dt]] = paste("buffer_", names(lists98)[dt], sep = "")
    }
metas98



meta11 <- metaMDS(multi_env1000_11_pro[3:9])
meta98 <- metaMDS(multi_env1000_98_pro[3:9])

plot(meta11)
text(meta11, display = "species")
ordisurf(meta11, multi_env1000_11_pro$neonic, add = T)
ordisurf(meta11, multi_env1000_11_pro$pesticide, 
         col = "darkblue", add = T)

plot(meta98)
text(meta98, display = "species")
ordisurf(meta98, multi_env1000_98_pro$neonic, add = T)
ordisurf(meta98, multi_env1000_98_pro$pesticide, 
         col = "darkblue", add = T)

plot(procrustes(meta11, meta98))


## multivariate pesticides #####

````

```{r}
scale_glm <- glm(data = BBS_ln_scale3,
                     sum_abundance ~ scale_neonics:NEONIC * AG_SCALED + factor(YEAR) , family = poisson)
summary(scale_glmer)
coefplot2(scale_glmer, intercept = T)

names(BBS_ln_scale3)
levels(BBS_ln_scale3$ROUTE_FAC)
class(BBS_ln_scale3$YEAR)
BBS_ln_scale3$YEAR_FAC <- as.factor(BBS_ln_scale3$YEAR)

BBS_ln_scale3$AG_SCALED <- as.numeric(scale(BBS_ln_scale3$AGRICULTURE))
BBS_ln_scale3$FOREST_SCALED <- as.numeric(scale(BBS_ln_scale3$FOREST))

scale_glmer <- glmer(data = BBS_ln_scale3,
                     sum_abundance ~ scale_neonics:NEONIC * AG_SCALED + (1|ROUTE_FAC) + (1|YEAR_FAC), family = poisson)
summary(scale_glmer)
coefplot2(scale_glmer, intercept = T)

scale_glmer2 <- glmer(data = BBS_ln_scale3,
                     sum_abundance ~ FOREST_SCALED + scale_neonics:NEONIC * AG_SCALED + (1|ROUTE_FAC) + (1|YEAR_FAC), family = poisson)
summary(scale_glmer2)
coefplot2(scale_glmer2, intercept = T)
```
Now for the bigger buffers
```{r}
unique(BBS_ln_scale2$BUFFER)
BBS_ln_scale4 <- BBS_ln_scale2[BBS_ln_scale2$BUFFER == 10000 , ]
BBS_ln_scale4$obs <- 1:nrow(BBS_ln_scale4)

unique(BBS_ln_scale4$NEONIC)
BBS_ln_scale4$NEONIC <- factor(BBS_ln_scale4$NEONIC,
                                  levels = c("pesticide", "neonic"))  # changes the levels
BBS_ln_scale4$NEONIC <- relevel(BBS_ln_scale4$NEONIC, ref = "pesticide") #changes the ereference level
names(BBS_ln_scale4)

BBS_ln_scale4$YEAR_FAC <- as.factor(BBS_ln_scale4$YEAR)

BBS_ln_scale4$AG_SCALED <- as.numeric(scale(BBS_ln_scale4$AGRICULTURE))
BBS_ln_scale4$FOREST_SCALED <- as.numeric(scale(BBS_ln_scale4$FOREST))


scale_glmer_ten <- glmer(data = BBS_ln_scale4,
                     sum_abundance ~ scale_neonics:NEONIC * AG_SCALED + (1|ROUTE_FAC) + (1|YEAR_FAC), family = poisson)
summary(scale_glmer_ten)
coefplot2(scale_glmer_ten, intercept = T)

scale_glmer2_ten <- glmer(data = BBS_ln_scale4,
                     sum_abundance ~ FOREST_SCALED + scale_neonics:NEONIC * AG_SCALED + (1|ROUTE_FAC) + (1|YEAR_FAC), family = poisson)
summary(scale_glmer2_ten)
coefplot2(scale_glmer2_ten, intercept = T)

names(BBS_ln_scale4)
levels(BBS_ln_scale4$CONSERVATION)
levels(BBS_ln_scale4$BEHAVIOR)
BBS_ln_scale4$BEHAVIOR2 <- 
  ifelse(BBS_ln_scale4$BEHAVIOR == "ground_forager", 
         "ground_forager", 
         ifelse(BBS_ln_scale4$BEHAVIOR %in% c("aerial_diver","aerial_forager", "hovering", "soaring"), 
                "aerial", "other"))


scale_glmer3_ten <- glmer(data = BBS_ln_scale4,
                     sum_abundance ~ FOREST_SCALED   + scale_neonics:NEONIC * AG_SCALED + (1|ROUTE_FAC) + (1|YEAR_FAC), family = poisson)
summary(scale_glmer3_ten)
coefplot2(scale_glmer3_ten)
```

now to see what happens for different groups::

```{r}
steepdeclinebirds <- read.csv("raw_data/commonbirds_steepdecline_traits.csv")
steepdeclinebirds <- subset(steepdeclinebirds, ohio_presence > 0)

unique(BBS_ln_scale4$COMMON_NAME) %in% (steepdeclinebirds$common_name)

steepdeclinebirds <- steepdeclinebirds[steepdeclinebirds$common_name  %in%  sort(unique(BBS_ln_scale4$COMMON_NAME)), ]
#####

steepdecline_bbs2<- subset(BBS_ln_scale4, COMMON_NAME %in% steepdeclinebirds$common_name)
names(steepdecline_bbs)
steepdecline_bbs2 <- steepdecline_bbs %>%
  group_by(YEAR, COMMON_NAME, NESTING, BEHAVIOR, DIET, DIET_VULN, DIET_INSECTS) %>%
  summarise(mean_abund = mean(sum_abundance),
            se_abund = sd(sum_abundance)/sqrt(length(sum_abundance)))

ggplot(steepdecline_bbs2, aes(x = YEAR, y = mean_abund)) +
  geom_line(aes(colour = COMMON_NAME)) +theme_bw() +
  facet_wrap(~ COMMON_NAME, scales = "free")



scale_glmer2_steep <- glmer(data = subset(BBS_ln_scale3, COMMON_NAME %in% c("Eastern Meadowlark", "Field Sparrow", "Common Grackle", "Chimney Swift", "Northern Bobwhite")),
                     sum_abundance ~ FOREST_SCALED + scale_neonics:NEONIC * AG_SCALED + YEAR:AG_SCALED + YEAR:scale_neonics:NEONIC+ (1|ROUTE_FAC) + (1|YEAR_FAC) + (1|COMMON_NAME), family = poisson)
summary(scale_glmer2_steep)
coefplot2(scale_glmer2_steep)

```


```{r}
scale_glmer <- glmer(data = BBS_ln_scale3,
                     sum_abundance ~ scale_scale_:NEONIC +
                       scale_scale_ + AG_SCALED + FOREST_SCALED +
                       year_scaled +
                       DIET_INSECTS + DIET_INSECTS:scale_scale_:NEONIC +
                       (1|ROUTE_FAC) + (1|COMMON_NAME), family = poisson)
summary(scale_glmer)
coefplot2(scale_glmer, intercept = T)

levels(BBS_ln_scale3$DIET)

BBS_ln_scale3 %>% 
  group_by(DIET) %>%
  summarise(count_sp = length(unique(COMMON_NAME)))

BBS_ln_scale3$DIET2 <- factor(ifelse(BBS_ln_scale3$DIET == "seeds", "seedeater", ifelse(BBS_ln_scale3$DIET == "omnivore", "omnivorous", "eatsother")))

# 
# save(BBS_ln_scale3, file = "BBS_ln_scales.Rdata")
# save(BBS_ln_scale2, file = "BBS_ln_scales2.Rdata")

names(BBS_ln_scale3)
````


```{r}

scale_glmA <- glm(data = BBS_ln_scale3,
                     sum_abundance ~ factor(YEAR) * FOREST_SCALED * AG_SCALED, family = poisson)

summary(scale_glmA)

scale_glmerA <- glmer(data = BBS_ln_scale3,
                     sum_abundance ~ factor(YEAR) * FOREST_SCALED * AG_SCALED +
                       (1|ROUTE_FAC), family = poisson)

summary(scale_glmerA)
```


````{r}
scale_glmer2 <- glmer(data = BBS_ln_scale3,
                     sum_abundance ~ scale_scale_:NEONIC +
                       scale_scale_ + AG_SCALED +
                       year_scaled +
                       DIET2 + DIET2:scale_scale_:NEONIC +
                       (1|ROUTE_FAC) + (1|COMMON_NAME), family = poisson)
summary(scale_glmer2)
levels(BBS_ln_scale3$DIET2)
coefplot2(scale_glmer2, intercept = T)

scale_glmer3 <- glmer(data = BBS_ln_scale3,
                     sum_abundance ~ scale_scale_:NEONIC ++ AG_SCALED + FOREST_SCALED +
                       year_scaled +
                       DIET2 + DIET2:scale_scale_:NEONIC +
                       (1|ROUTE_FAC) , family = poisson)
summary(scale_glmer3)
coefplot2(scale_glmer3, intercept = T)

scale_glmer4 <- glmer(data = BBS_ln_scale3,
                     sum_abundance ~ scale_scale_:NEONIC + AG_SCALED + AG_SCALED:scale_scale_ + FOREST_SCALED +
                       year_scaled +
                       DIET2 + DIET2:scale_scale_:NEONIC +
                       (1|ROUTE_FAC) + (1|COMMON_NAME), family = poisson)
summary(scale_glmer4)
coefplot2(scale_glmer3, intercept = T)

scale_glmer5 <- glmer(data = BBS_ln_scale3,
                     sum_abundance ~ scale_scale_:NEONIC + AG_SCALED + AG_SCALED:scale_scale_ + 
                       year_scaled +
                       DIET2 + DIET2:scale_scale_:NEONIC +DIET2:scale_scale_:NEONIC:AG_SCALED +
                       (1|ROUTE_FAC) + (1|COMMON_NAME), family = poisson)
summary(scale_glmer5)

anova(scale_glmer2, scale_glmer3)



# test_mod2 <- glmer(data = BBS_ln[ BBS_ln$buffer == 200, ],
#                    family = poisson,
#                    formula= SUM_ROUTE_ABUNDANCE ~ agriculture_scaled + high_kg_buff_scaled:COMPOUND + DIET + (1|BBS_route) + (1|YEAR))
#summary(test_mod2)
require(coefplot2)
#coefplot2(test_mod2)


require(dismo)
BBS_200 <- BBS_ln[ BBS_ln$buffer == 200, ]
names(BBS_200)
class(BBS_200$SPECIES)
# gbm_neonic <- gbm.step(data = BBS_200, 
#                      gbm.x=c(1, 3, 15, 16, 17, 18, 23, 25, 26, 29:35), 
#                      gbm.y = 11, family = "poisson",
#                      tree.complexity = 14) #, 
#                       #step.size = 20, n.trees = 50,
#                      #learning.rate = 0.001, bag.fraction = 0.75)
# summary(gbm_neonic)
# 
# glmm_neonic <- glmer(data = BBS_200,
#                      formula = SUM_ROUTE_ABUNDANCE ~ high_kg_buff:COMPOUND + 
#                        high_kg_buff:forest:COMPOUND + 
#                        high_kg_buff:agriculture:COMPOUND + 
#                        forest + agriculture +
#                        (1|GENUS) + (1|YEAR),
#                      family = "poisson")
# 
# summary(glmm_neonic)
# coefplot2(glmm_neonic)
# 
# neonic_int<-gbm.interactions(gbm_neonic)
# neonic_int$interactions
# 
# gbm.perspec(gbm_neonic, 1,3)
# gbm.perspec(gbm_neonic, 2,3)
# gbm.perspec(gbm_neonic, 4,8)
# gbm.plot(gbm_neonic, 3,4)
# 
# save.image(file = "big_gbm.Rdata")
# load(file = "big_gbm.Rdata")
#saveRDS(gbm_neonic, file = "big_serialised_gbm.rds")

```

# Some exploration! 
```{r, cache=TRUE}
head(landuse_buffers)

ggplot(landuse_buffers[landuse_buffers$land_cover %in% c("forest", "agriculture", "developed", "grassland_shrub") & landuse_buffers$buffer == 400, ], aes(x = YEAR, y = total_pix)) +
  geom_smooth(aes(colour = land_cover)) +
#  geom_density2d(aes(colour = land_cover)) +
  geom_point(aes(colour = land_cover), alpha = 0.3, position = position_jitter(width = 0.3))+
  theme_bw()+
  scale_colour_colorblind("Land Cover", labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub"))+
  guides(colour=guide_legend(override.aes=list(fill=NA)))+
  theme(legend.position = "bottom", legend.key = element_blank(), legend.background = element_rect(colour = "grey"))

ggplot(landuse_buffers[landuse_buffers$land_cover %in% c("forest", "agriculture", "developed", "grassland_shrub") & landuse_buffers$buffer == 400, ], aes(x = YEAR, y = total_pix)) +
  geom_smooth(aes(colour = land_cover)) +
  geom_density2d(aes(colour = land_cover)) +
  #geom_point(aes(colour = land_cover), alpha = 0.3, position = position_jitter(width = 0.3))+
  theme_bw()+
  scale_colour_colorblind("Land Cover", labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub"))+
  guides(colour = FALSE) +
  facet_wrap(~ land_cover, scales = "free")

ggplot(landuse_buffers[landuse_buffers$land_cover %in% c("forest", "agriculture", "developed", "grassland_shrub") & landuse_buffers$buffer == 400 & landuse_buffers$YEAR %in% c(1995, 2000, 2005, 2010), ], aes(x = land_cover, y = total_pix)) +
 geom_violin(aes(fill = land_cover))+
  theme_bw()+
  scale_fill_colorblind("Land Cover", labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
  facet_wrap(~YEAR)


ggplot(landuse_buffers[landuse_buffers$land_cover %in% c("forest", "agriculture", "developed", "grassland_shrub") & landuse_buffers$buffer == 400 & landuse_buffers$YEAR %in% c(1995, 2000, 2005, 2010), ], aes(x = factor(YEAR), y = total_pix)) +
 geom_violin(aes(fill = land_cover))+
  theme_bw()+
  scale_fill_colorblind("Land Cover", labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
  guides(fill = FALSE)+
  facet_wrap(~land_cover, scales = "free")


summary_landuse <- landuse_buffers[landuse_buffers$land_cover %in% c("forest", "agriculture", "developed", "grassland_shrub"), ] %>% group_by(YEAR, land_cover, buffer) %>% filter(buffer == 400) %>% summarise(total_pix = sum(total_pix))

ggplot(summary_landuse, aes(x = YEAR, y = total_pix)) +
  geom_area(aes(fill = land_cover, colour = land_cover), position = "fill")+
  theme_bw()+
  scale_fill_manual("Land Cover", values = c("sienna4", "gray20", "forestgreen", "orange"), labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
  scale_colour_manual("Land Cover", values = c("sienna4", "gray20", "forestgreen", "orange"), labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
  theme(legend.position = "bottom", legend.background = element_rect(colour = "grey"))

unique(landuse_buffers$buffer)
summary_landuse3 <- landuse_buffers[landuse_buffers$land_cover %in% c("forest", "agriculture", "developed", "grassland_shrub"), ] %>% group_by(YEAR, land_cover, buffer) %>% summarise(total_pix = sum(total_pix))

ggplot(summary_landuse3, aes(x = YEAR, y = total_pix)) +
  geom_area(aes(fill = land_cover, colour = land_cover), position = "fill")+
  theme_bw()+
  scale_fill_manual("Land Cover", values = c("sienna4", "gray20", "forestgreen", "orange"), labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
  scale_colour_manual("Land Cover", values = c("sienna4", "gray20", "forestgreen", "orange"), labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
  theme(legend.position = "bottom", legend.background = element_rect(colour = "grey")) +
  facet_wrap(~buffer)


# summary_landuse2 <- landuse_buffers%>% group_by(YEAR, land_cover, buffer) %>% filter(buffer == 400) %>% summarise(total_pix = sum(total_pix))
# 
# ggplot(summary_landuse2, aes(x = YEAR, y = total_pix)) +
#   geom_area(aes(fill = land_cover, colour = land_cover), position = "fill")+
#   theme_bw()+
#   #scale_fill_manual("Land Cover", values = c("sienna4", "gray20", "forestgreen", "orange"), labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
#   #scale_colour_manual("Land Cover", values = c("sienna4", "gray20", "forestgreen", "orange"), labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
#   theme(legend.position = "bottom", legend.background = element_rect(colour = "grey"))

````

# Exploring the neonics

```{r, cache=TRUE}
head(neonics_buffers)

head(neonics_buffers[is.na(neonics_buffers$high_kg_buff) | is.na(neonics_buffers$ag_pix_buff) | is.na(neonics_buffers$low_kg_buff), ])
head(neonics_buffers[is.na(neonics_buffers$high_kg_buff) | is.na(neonics_buffers$ag_pix_buff) , ])
dim(neonics_buffers[is.na(neonics_buffers$high_kg_buff) | is.na(neonics_buffers$ag_pix_buff) | is.na(neonics_buffers$low_kg_buff), ])
dim(neonics_buffers[is.na(neonics_buffers$high_kg_buff) | is.na(neonics_buffers$ag_pix_buff), ])

# unique(neonics_buffers$YEAR)
# class(neonics_buffers$high_kg_buff)
# noyear_neonics_buffers <- neonics_buffers %>% 
#   group_by(BBS_route, buffer) %>% 
#   summarise(high_pest = sum(high_kg_buff), agpix = sum(ag_pix_buff))
# head(noyear_neonics_buffers)
# # spps <- spps[!is.na(spps$NMDS1) & !is.na(spps$NMDS2),]
# noyear_neonics_buffers <- noyear_neonics_buffers[!is.na(noyear_neonics_buffers$high_pest) & !is.na(noyear_neonics_buffers$agpix), ]
# 
# ggplot(noyear_neonics_buffers, aes(y = as.numeric(high_pest/agpix), x = as.numeric(buffer))) +
#   geom_line(aes(group = BBS_route)) +
#   geom_smooth()+
#   theme_bw()
# 
# ggplot(noyear_neonics_buffers, aes(y = high_pest, x = buffer)) +
#   geom_line(aes(group = BBS_route)) +
#   geom_smooth()+
#   theme_bw()
# 
# ggplot(noyear_neonics_buffers, aes(y = agpix, x = buffer)) +
#   geom_line(aes(group = BBS_route)) +
#   geom_smooth()+
#   theme_bw()
# 
# 
# # now to see if intensity has increased over tine
# names(neonics_buffers)
# year_neonics_buffers <- neonics_buffers %>%
#   group_by(BBS_route, YEAR, COMPOUND, buffer) %>%
#   summarise(intensity = high_kg_buff/ag_pix_buff) 
# 
# names(year_neonics_buffers)
# 
# ggplot(year_neonics_buffers, aes(y = intensity, x = YEAR)) +
#   geom_line(aes(group = BBS_route)) +
#   facet_grid(buffer~ COMPOUND, scale = "free") +
#   theme_bw()
# 
# year_neonics_buffers2 <- neonics_buffers %>%
#   group_by(BBS_route, YEAR, COMPOUND) %>%
#   summarise(intensity = mean(high_kg_buff/ag_pix_buff)) 
# 
# ggplot(year_neonics_buffers2, aes(y = intensity, x = YEAR)) + 
#   geom_line(aes(group = BBS_route)) +
#   facet_wrap(~ COMPOUND, scale = "free") +
#   theme_bw()
# 
# ggplot(year_neonics_buffers2, aes(y = intensity, x = YEAR)) + 
#   geom_line(aes(group = BBS_route)) +
#   facet_wrap(~ COMPOUND) +
#   theme_bw()

```

# Using the big dataset: BBS_ln
```{r}
head(BBS_ln)
names(BBS_ln) <- toupper(names(BBS_ln))
sort(unique(BBS_ln$DIET))
BBS_ln$DIET <- relevel(BBS_ln$DIET, ref = "insects")
BBS_ln$YEAR_SCALED <- scale(BBS_ln$YEAR)
#BBS_ln$FOREST_SCALED <- scale(BBS_ln$FOREST)
#BBS_ln$AGRICULTURE_SCALED <- scale(BBS_ln$AGRICULTURE)
#BBS_ln$FOREST_PROP <- BBS_ln$FOREST/BBS_ln$AGRICULTURE
#BBS_ln$HIGH_KG_BUFF <- scale(BBS_ln$HIGH_KG_BUFF)


with(BBS_ln, cor(FOREST, AGRICULTURE))
# test_mod <- glm(SUM_ROUTE_ABUNDANCE ~ YEAR_SCALED + 
#                   AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
#                   AGRICULTURE_SCALED +
#                   DIET_VULN:YEAR_SCALED + 
#                   DIET_VULN:HIGH_KG_BUFF_SCALED:YEAR_SCALED:AGRICULTURE_SCALED + 
#                   HIGH_KG_BUFF_SCALED:COMPOUND + 
#                   HIGH_KG_BUFF_SCALED, 
#                 data = BBS_ln[BBS_ln$COMPOUND %in% c("ATRAZINE", "GLYPHOSATE"),], 
#                 family = "poisson")
# summary(test_mod)
# #coefplot2(test_mod)
# #require(effects)
# 
# 
# test_mod2 <- glm(SUM_ROUTE_ABUNDANCE ~ YEAR_SCALED + 
#                   AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
#                   AGRICULTURE_SCALED +
#                   DIET_INSECTS:YEAR_SCALED + 
#                   DIET_INSECTS:HIGH_KG_BUFF_SCALED:YEAR_SCALED:AGRICULTURE_SCALED + 
#                   HIGH_KG_BUFF_SCALED:COMPOUND + 
#                   HIGH_KG_BUFF_SCALED, 
#                 data = BBS_ln[BBS_ln$COMPOUND %in% c("ATRAZINE", "GLYPHOSATE"),], 
#                 family = "poisson")
# summary(test_mod2)
# #coefplot2(test_mod2)
# 
# 
# glmer_mod_3 <- glmer(SUM_ROUTE_ABUNDANCE ~ YEAR_SCALED + 
#                   AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
#                   AGRICULTURE_SCALED +
#                   DIET_INSECTS:YEAR_SCALED + 
#                   DIET_INSECTS:HIGH_KG_BUFF_SCALED:YEAR_SCALED:AGRICULTURE_SCALED + 
#                   HIGH_KG_BUFF_SCALED:COMPOUND + 
#                   HIGH_KG_BUFF_SCALED +
#                     (1|ROUTE), 
#                 data = BBS_ln[BBS_ln$COMPOUND %in% c("ATRAZINE", "GLYPHOSATE"),], 
#                 family = "poisson")
# #coefplot2(glmer_mod_3)
# #coefplot2(list(glmer_mod_3, test_mod2))
# 
# glmer_mod_4 <- glmer(SUM_ROUTE_ABUNDANCE ~ YEAR_SCALED + 
#                   AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
#                   AGRICULTURE_SCALED +
#                   DIET_INSECTS:YEAR_SCALED + 
#                     DIET_INSECTS +
#                   DIET_INSECTS:HIGH_KG_BUFF_SCALED:YEAR_SCALED:AGRICULTURE_SCALED + 
#                   HIGH_KG_BUFF_SCALED:COMPOUND + 
#                   HIGH_KG_BUFF_SCALED +
#                     (1|ROUTE), 
#                 data = BBS_ln[BBS_ln$COMPOUND %in% c("ATRAZINE", "GLYPHOSATE"),], 
#                 family = "poisson")
# summary(glmer_mod_4)
# #coefplot2(glmer_mod_4)
# #coefplot2(list(glmer_mod_3, test_mod2, glmer_mod_4))
# 
# glmer_mod_5 <- glmer(SUM_ROUTE_ABUNDANCE ~ 
#                        AGRICULTURE_SCALED:DIET_INSECTS + 
#                        AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
#                        FOREST_SCALED:DIET_INSECTS +
#                        DIET_INSECTS:HIGH_KG_BUFF_SCALED:AGRICULTURE_SCALED + 
#                        HIGH_KG_BUFF_SCALED:COMPOUND:DIET_INSECTS + 
#                        HIGH_KG_BUFF_SCALED:DIET_INSECTS +
#                        (ROUTE|YEAR), 
#                 data = BBS_ln[BBS_ln$COMPOUND %in% c("ATRAZINE", "GLYPHOSATE"),], 
#                 family = "poisson")
# summary(glmer_mod_5)
#coefplot2(glmer_mod_5)
#coefplot2(list(glmer_mod_3, test_mod2, glmer_mod_4, glmer_mod_5))
```

# Aggregating based on diet and then modelling

```{r}
head(BBS_ln)
names(BBS_ln)
# BBS_ln$HIGH_KG_BUFF <- as.numeric(BBS_ln$HIGH_KG_BUFF)
# BBS_ln$AGRICULTURE_SCALED <- as.numeric(BBS_ln$AGRICULTURE_SCALED)
# BBS_ln$HIGH_KG_BUFF_SCALED <- as.numeric(BBS_ln$HIGH_KG_BUFF_SCALED)
# BBS_ln$YEAR_SCALED <- as.numeric(BBS_ln$YEAR_SCALED)
# BBS_ln$FOREST_SCALED <- as.numeric(BBS_ln$FOREST_SCALED)
# BBS_ln_diet <- data.frame(BBS_ln %>% group_by(YEAR, RTENO, BBS_ROUTE, BUFFER, ROUTE, COMPOUND, HIGH_KG_BUFF, AGRICULTURE, FOREST, DEVELOPED, AGRICULTURE_SCALED, YEAR_SCALED, FOREST_SCALED, HIGH_KG_BUFF_SCALED, FOREST_PROP, DIET) %>%
#   summarise(SUM_ROUTE_ABUNDANCE_DIET = sum(SUM_ROUTE_ABUNDANCE)))
# head(BBS_ln_diet)

````

```{r}
class(BBS_ln_diet$ROUTE)
BBS_ln_diet$YEAR_fac <- as.factor(BBS_ln_diet$YEAR)
BBS_ln_diet$ROUTE <- as.factor(BBS_ln_diet$ROUTE)


range(na.omit(BBS_ln_diet$HIGH_KG_BUFF_SCALED))
range(na.omit(BBS_ln_diet$AGRICULTURE_SCALED))
range(na.omit(BBS_ln_diet$FOREST_SCALED))
range(na.omit(BBS_ln_diet$FOREST_SCALED))
class(BBS_ln_diet$DIET)
class(BBS_ln_diet$ROUTE)
range(BBS_ln_diet$SUM_ROUTE_ABUNDANCE_DIET)
ggplot(BBS_ln_diet, aes(x = ROUTE, y = SUM_ROUTE_ABUNDANCE_DIET)) +
  geom_point() +
  geom_boxplot() + theme_bw() +
  facet_wrap(~ YEAR)

BBS_ln_diet$sqrt_response <- round(sqrt(BBS_ln_diet$SUM_ROUTE_ABUNDANCE_DIET), 0)
range(BBS_ln_diet$sqrt_response)

BBS_ln_dietcc <- BBS_ln_diet[complete.cases(BBS_ln_diet),]
str(BBS_ln_dietcc)

# load("predicting_glmers.Rdata")
require(lme4)
# glmer_mod_6 <- glmer(na.action = na.omit,
#                      sqrt_response ~ 
#                        AGRICULTURE_SCALED:DIET+ 
#                        AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
#                        FOREST_SCALED:DIET +
#                        DIET:HIGH_KG_BUFF_SCALED:AGRICULTURE_SCALED + 
#                        HIGH_KG_BUFF_SCALED:COMPOUND:DIET + 
#                        HIGH_KG_BUFF_SCALED:DIET +
#                        (1|YEAR_fac) + (1|ROUTE), 
#                      data = BBS_ln_dietcc[BBS_ln_dietcc$COMPOUND %in% c("ATRAZINE", 
#                                                                         "GLYPHOSATE")
#                                           & BBS_ln_dietcc$YEAR > 1996
#                                           & BBS_ln_dietcc$DIET != "",], 
#                      family = "poisson", 
#                      verbose = TRUE)
# sum_glmer_mod_6<-summary(glmer_mod_6)
# cf2_glmer_mod_6 <- coefplot2(glmer_mod_6)
# save(glmer_mod_6, sum_glmer_mod_6, cf2_glmer_mod_6, file = "glmer_mod_6.Rdata")
# load( "glmer_mod_6.Rdata")
# save.image("predicting_glmers.Rdata")
summary(glmer_mod_6)
diet_levels <- levels(BBS_ln_dietcc$DIET)[-2]
ep <- expand.grid(AGRICULTURE_SCALED = seq(from = min(BBS_ln_dietcc$AGRICULTURE_SCALED), 
                                    to = max(BBS_ln_dietcc$AGRICULTURE_SCALED),
                                    by = 0.4),
                  FOREST_SCALED = mean(BBS_ln_dietcc$FOREST_SCALED),
                  DIET = diet_levels, 
                  HIGH_KG_BUFF_SCALED =seq(from = min(BBS_ln_dietcc$HIGH_KG_BUFF_SCALED), 
                                    to = max(BBS_ln_dietcc$HIGH_KG_BUFF_SCALED),
                                    by = 2),
                  COMPOUND = c("GLYPHOSATE", "ATRAZINE")
                    )

mm <- model.matrix(delete.response(terms(glmer_mod_6)), ep)

ep <- data.frame(ep, response = mm %*% fixef(glmer_mod_6))
ep <- with(ep, data.frame(ep, transformed.response = exp(response)))

pvar1 <- diag(mm %*% tcrossprod(vcov(glmer_mod_6), mm))
tvar1 <- pvar1 + VarCorr(glmer_mod_6)$YEAR + VarCorr(glmer_mod_6)$ROUTE
ep <- data.frame(ep, pse=sqrt(pvar1), tse=sqrt(tvar1))
ep <- with(ep,
                data.frame(ep,
                           plo=exp(response-1.96*pse),
                           phi=exp(response+1.96*pse),
                           tlo=exp(response-1.96*tse),
                           thi=exp(response+1.96*tse)))
head(ep)

names(ep) <- tolower(names(ep))
names(ep) 
ep1 <- ep[ep$diet %in% c("insects", "omnivore"), ]
require(akima)
ep2 <- with(ep1[ep1$compound == "GLYPHOSATE" & ep1$diet == "insects",], interp(agriculture_scaled, high_kg_buff_scaled, transformed.response))
ep3 <- with(ep1[ep1$compound == "GLYPHOSATE" & ep1$diet == "omnivore",], interp(agriculture_scaled, high_kg_buff_scaled, transformed.response))

ep4 <- with(ep1[ep1$compound == "ATRAZINE" & ep1$diet == "insects",], interp(agriculture_scaled, high_kg_buff_scaled, transformed.response))
ep5 <- with(ep1[ep1$compound == "ATRAZINE" & ep1$diet == "omnivore",], interp(agriculture_scaled, high_kg_buff_scaled, transformed.response))


with(ep2, persp(x, y, z, theta = 330, phi = 20, 
                xlab = "Agriculture",
                ylab = "Pesticide application rate", 
                zlab = "Bird response",
                col = "royalblue",
                #nticks = 5, 
                #ticktype = "detailed",
                shade = 0.75))

with(ep3, persp(x, y, z, theta = 55, phi = 30, 
                xlab = "Agriculture",
                ylab = "Pesticide application rate", 
                zlab = "Bird response",
                col = "cornflowerblue",
                #nticks = 5, 
                #ticktype = "detailed",
                shade = 0.75))

with(ep4, persp(x, y, z, theta = 280, phi = 30, 
                xlab = "Agriculture",
                ylab = "Pesticide application rate", 
                zlab = "Bird response",
                col = "orange",
                #nticks = 5, 
                #ticktype = "detailed",
                shade = 0.75))

with(ep5, persp(x, y, z, theta = 55, phi = 30, 
                xlab = "Agriculture",
                ylab = "Pesticide application rate", 
                zlab = "Bird response",
                col = "violetred",
                #nticks = 5, 
                #ticktype = "detailed",
                shade = 0.2))


```

Earlier glm model which you can run if interested, heavily pseudo-replicated however.
```{r}
glm_mod_6 <- glm(na.action = na.omit,
                 sqrt_response ~ 
                   DIET+
                   HIGH_KG_BUFF_SCALED +
                   AGRICULTURE_SCALED:DIET+ 
                   AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
                   DIET:HIGH_KG_BUFF_SCALED:AGRICULTURE_SCALED + 
                   HIGH_KG_BUFF_SCALED:COMPOUND:DIET + 
                   HIGH_KG_BUFF_SCALED:DIET, 
                 data = BBS_ln_dietcc[BBS_ln_dietcc$COMPOUND %in% c("ATRAZINE", 
                                                                    "GLYPHOSATE")
                                      & BBS_ln_dietcc$YEAR > 1996
                                      & BBS_ln_dietcc$DIET != "",], 
                 family = "poisson")
summary(glm_mod_6)
coefplot2(glm_mod_6)

```