---
title: "BBS_neonics_NMDS"
author: "Olivia Burge"
date: "Wednesday, December 03, 2014"
output: html_document
---

```{r packagesetup}

#best you be changing the repos. Whatever repos you usually use isn't picked up 
#when knitting and you need to install and it just throws an error.
if(!require(ggplot2)){
    install.packages("ggplot2", repos = 'http://cran.stat.auckland.ac.nz/')
    library(ggplot2)
 }
if(!require(lme4)){
    install.packages("lme4", repos = 'http://cran.stat.auckland.ac.nz/')
    library(lme4)
}
if(!require(tidyr)){
    install.packages("tidyr", repos = 'http://cran.stat.auckland.ac.nz/')
    library(tidyr)
}
if(!require(stringr)){
    install.packages("stringr", repos = 'http://cran.stat.auckland.ac.nz/')
    library(stringr)
}
if(!require(grid)){
  install.packages("grid", repos = 'http://cran.stat.auckland.ac.nz/')
  library(grid)
}
if(!require(RCurl)){
  install.packages("RCurl", repos = 'http://cran.stat.auckland.ac.nz/')
  library(RCurl)
}
if(!require(dplyr)){
  install.packages("dplyr", repos = 'http://cran.stat.auckland.ac.nz/')
  library(dplyr)
}

if(!require(ggthemes)){
  install.packages("ggthemes", repos = 'http://cran.stat.auckland.ac.nz/')
  library(ggthemes)
}
if(!require(vegan)){
  install.packages("vegan", repos = 'http://cran.stat.auckland.ac.nz/')
  library(vegan)
}
````

```

# AOU Codes
Required because BBS uses "AOU" codes for its species, but they aren't the same as everywhere.  These ones are sourced from their website, and post-processed into a csv.
````{r, AOUdata, results = "hide"}
#setwd("/Users/Liv/Documents/NCEAS_GIT/NCEAS-RENCI_2014/BBS_data")
AOU_codes <- read.csv("raw_data/AOU_codes.csv") ## a cleaned up (ie header removed, split into columns - no substantive changes to any species names or column headers) version of: ftp://ftpext.usgs.gov/pub/er/md/laurel/BBS/DataFiles/SpeciesList.txt
head(AOU_codes)
AOU_codes <- AOU_codes[ - 4] # removing the French language names which have a bunch of mis-translated characters in 
names(AOU_codes) <- toupper(names(AOU_codes))
````

# BBS data
I got the BBS data from: ftp://ftpext.usgs.gov/pub/er/md/laurel/BBS/DataFiles//50-StopData/1997ToPresent_SurveyWide/Fifty7.zip.  The metadata is available: ftp://ftpext.usgs.gov/pub/er/md/laurel/BBS/DataFiles//FiftySt.txt.

````{r, results = "hide"}
ohio_BBS <- read.csv("raw_data/fifty7.csv")
names(ohio_BBS) <- toupper(names(ohio_BBS))
head(ohio_BBS)

#adding abundance
ohio_BBS$ROUTE_ABUNDANCE <- rowSums(ohio_BBS[grep("^STOP[0-9]+", names(ohio_BBS))])
head(ohio_BBS)
ohio_BBS <- ohio_BBS %>% select(ROUTE,  YEAR, AOU, ROUTE_ABUNDANCE)
head(ohio_BBS)


names(ohio_BBS)
names(AOU_codes)
## merging in the species names
ohio_BBS_AOU <-inner_join(ohio_BBS, AOU_codes)

#now we need to locate the routes in each county.  I read in the file Tyson had already located for us:
ohio_BBS_routes_countiesraw <- read.csv("raw_data/BBSdata.by.county.csv")
names(ohio_BBS_routes_countiesraw) <- toupper(names(ohio_BBS_routes_countiesraw))
head(ohio_BBS_routes_countiesraw)
ohio_BBS_routes_countiesraw2 <- ohio_BBS_routes_countiesraw[!duplicated(ohio_BBS_routes_countiesraw$ROUTE), c(1,4,5,6,7)]

names(ohio_BBS_AOU) %in% names(ohio_BBS_routes_countiesraw2)
#and then joined the two
ohio_BBS_birds <- inner_join(ohio_BBS_AOU, ohio_BBS_routes_countiesraw2)
head(ohio_BBS_birds)
names(ohio_BBS_birds)

```

```{r actualNMDS}

## data prep
head(ohio_BBS_birds)
names(ohio_BBS_birds) <- toupper(names(ohio_BBS_birds))
names(ohio_BBS_birds) 

# need one value per route per year
ohio_BBS_birds_shortB <- data.frame(ohio_BBS_birds %>% group_by(ROUTE, COUNTY, GEOID, YEAR, COMMON_NAME)  %>% filter(YEAR >1996 & YEAR < 2010) %>% summarise(ROUTE_ABUNDANCE = mean(ROUTE_ABUNDANCE)))
head(ohio_BBS_birds_shortB)

# need a matrix with species as columns
ohio_BBS_birds_short <- ohio_BBS_birds_shortB %>% spread(key = COMMON_NAME, value = ROUTE_ABUNDANCE, fill = 0)

names(ohio_BBS_birds_short)
head(ohio_BBS_birds_short)
set.seed(333) # reproducibility

# NMDS
nmds_ohio <- metaMDS(ohio_BBS_birds_short[ , 5:313])

# quick visual
plot(nmds_ohio)
ordiellipse(nmds_ohio, ohio_BBS_birds_short$YEAR)

# envfit - by year and county
env_birds <- envfit(nmds_ohio, ohio_BBS_birds_short[,c(2,4)])

# adonis - effefct of year after accounting for county differences, strata by route. + interaction effct
adonis_birds <- adonis(ohio_BBS_birds_short[, 5:313] ~ COUNTY * YEAR, strata = ohio_BBS_birds_short$ROUTE, data =ohio_BBS_birds_short )

# betadisper
birds_dist <- vegdist(ohio_BBS_birds_short[, 5:313], method = "bray")
beta_nenonic <- betadisper(birds_dist, ohio_BBS_birds_short$YEAR)
beta_nenonic

boxplot(beta_nenonic, ylab = "Distance to centroid")
TukeyHSD(beta_nenonic, ordered = TRUE)

## plot

birdMDSdata <- data.frame(ROUTE = ohio_BBS_birds_short$ROUTE, 
                          GEOID = ohio_BBS_birds_short$GEOID,
                          COUNTY = ohio_BBS_birds_short$COUNTY,
                          YEAR = ohio_BBS_birds_short$YEAR,
                          NMDS1 = nmds_ohio$points[,1],
                          NMDS2 = nmds_ohio$points[,2])

## Birds NMDS.. obivously need to get rid of 1996.
birdplot_1 <- ggplot(birdMDSdata, aes (x = NMDS1, y = NMDS2)) +
  geom_point(aes(colour = factor(YEAR))) +
  geom_density2d(aes(colour = factor(YEAR), group = YEAR)) +
  facet_wrap(~ YEAR) +
  scale_colour_discrete(guide = FALSE)+
  theme_bw() 


birdplot_1

````

# Why? 
This is all just practice.  Need to add in feeding guilds to the above and the neonics/pesticides.

```{r neonics}
## come back to here!
# neonics_raw_ohioj <- spread(neonics_raw_ohio, key = COMPOUND, value = KG) # not working
# dim(neonics_raw_ohioj)
# dim(neonics_raw_ohio)
# dim(ohio_BBS_birds_short2) #merge by rows, exclude the right years.
# neonvegbirds <- merge(neonics_raw_ohio, ohio_BBS_birds_short2)





```