---
title: 'Ohio: BBS bird abundance'
author: "Olivia Burge"
date: "5 October 2014"
output: pdf_document
---

# Task one: the traits
I've saved a csv of common birds in steep decline from Kelly's xlsx.  The only change I made in excel was deleting the apostrophes as they looked _nasty_ when read in.  Otherwise, nada.  I haven't set a working directory for this and used full paths because I like that.  But if you guys give me a stern talking to I will.  Slash I will once I put all this stuffs in one place.

```{r, include = FALSE}
#best you be changing the repos. Whatever repos you usually use isn't picked up 
#when knitting and you need to install and it just throws an error.
if(!require(ggplot2)){
    install.packages("ggplot2", repos = 'http://cran.stat.auckland.ac.nz/')
    library(ggplot2)
}
if(!require(lme4)){
    install.packages("lme4", repos = 'http://cran.stat.auckland.ac.nz/')
    library(lme4)
}
if(!require(tidyr)){
    install.packages("tidyr", repos = 'http://cran.stat.auckland.ac.nz/')
    library(tidyr)
}
if(!require(stringr)){
    install.packages("stringr", repos = 'http://cran.stat.auckland.ac.nz/')
    library(stringr)
}
if(!require(grid)){
  install.packages("grid", repos = 'http://cran.stat.auckland.ac.nz/')
  library(grid)
}
if(!require(RCurl)){
  install.packages("RCurl", repos = 'http://cran.stat.auckland.ac.nz/')
  library(RCurl)
}
if(!require(dplyr)){
  install.packages("dplyr", repos = 'http://cran.stat.auckland.ac.nz/')
  library(dplyr)
}

if(!require(ggthemes)){
  install.packages("ggthemes", repos = 'http://cran.stat.auckland.ac.nz/')
  library(ggthemes)
}



````

````{r, results = "hide", cache=FALSE}
setwd("/Users/Liv/Documents/NCEAS_GIT/NCEAS-RENCI_2014/BBS_data")
traits_commmon_steep <- read.csv("raw_data/commonbirds_steepdecline_traits.csv")
str(traits_commmon_steep)

#subsetting just the birds that are in Ohio and eat insect
traits_commmon_steep_insectohio <- as.data.frame(traits_commmon_steep %>%
  filter(ohio_presence > 0 & diet == "insects"))

str(traits_commmon_steep_insectohio)

#because we use it later
traits_commmon_steep_insectohio$common_name <- factor(toupper(traits_commmon_steep_insectohio$common_name))

#which birds are we looking at? :
levels(traits_commmon_steep_insectohio$common_name)
```

#Task two - AOU codes for bird species

Next I also had to import the AOU codes because those are used in BBS, so we have a common 'key'. 

````{r, results = "hide"}
AOU_codes <- read.csv("raw_data/AOU_codes.csv")

# merge the AOU codes into the ohio data
traits_commmon_steep_insectohio2 <- as.data.frame(inner_join(AOU_codes, traits_commmon_steep_insectohio))

head(traits_commmon_steep_insectohio2)
````

# Task three: BBS data
I got the BBS data from: ftp://ftpext.usgs.gov/pub/er/md/laurel/BBS/DataFiles//50-StopData/1997ToPresent_SurveyWide/Fifty7.zip.  The metadata is available: ftp://ftpext.usgs.gov/pub/er/md/laurel/BBS/DataFiles//FiftySt.txt.

````{r, results = "hide"}
ohio_BBS <- read.csv("raw_data/fifty7.csv")
head(ohio_BBS)

# merge with the traits data
ohio_BBS_insect <- as.data.frame(inner_join(ohio_BBS, traits_commmon_steep_insectohio2))
ohio_BBS_insect$common_name<-factor(ohio_BBS_insect$common_name)
ohio_BBS_insect$diet <- factor(ohio_BBS_insect$diet)

#now we need to locate the routes in each county.  I read in the file Tyson had already located for us:
ohio_BBS_routes_countiesraw <- read.csv("raw_data/BBSdata.by.county.csv")
head(ohio_BBS_routes_countiesraw)
ohio_BBS_routes_countiesraw2 <- ohio_BBS_routes_countiesraw[!duplicated(ohio_BBS_routes_countiesraw$Route), c(1,4,5,6,7)]
 
#and then joined the two
ohio_BBS_birds <- inner_join(ohio_BBS_insect, ohio_BBS_routes_countiesraw2)
head(ohio_BBS_birds)
names(ohio_BBS_birds)

#ohio all BBS birds geolocated
ohio_all_birds_geo <- inner_join(ohio_BBS, ohio_BBS_routes_countiesraw2)
#anddd summed birds for each route:
ohio_all_birds_geo$route_abundance <- rowSums(ohio_all_birds_geo[grep("^Stop[0-9]+", names(ohio_all_birds_geo))])
names(ohio_all_birds_geo)
ohio_all_birds_geo <- ohio_all_birds_geo[ , c(1:7, 58:62)]
head(ohio_all_birds_geo)

# created a column of sums for all the stops on each route:
ohio_BBS_birds$route_abundance <- rowSums(ohio_BBS_birds[grep("^Stop[0-9]+", names(ohio_BBS_birds))])
#using grep means even if the columns change position, we will still get the right ones for rowsums.

```

Next up: a summary.  Some counties have more than one route.  Thought on this? 

# Getting a version of the neonics

```{r, results = "hide", cache=FALSE}
#this is from Tyson's work.
filenames <- list.files(path = "raw_data/raw_pesticides/", recursive = TRUE, full.names = TRUE) 
#I've amended the line above so that we can get the file paths as well 
#(the recursive and full.names options).

pesticides_raw <- do.call("rbind", lapply(filenames[grep("high", filenames)], 
                                          read.table, header = TRUE, sep = "\t"))

pesticides_raw_ohio <- pesticides_raw[which(pesticides_raw$STATE_FIPS_CODE == 39), ]

rm(pesticides_raw) #cos it's massive

names(pesticides_raw_ohio)

#getting the neonics out
neonics_raw_ohio <- pesticides_raw_ohio %>% 
  filter(COMPOUND %in% c("ACETAMIPRID", "CLOTHIANIDIN", "IMIDACLOPRID", "THIAMETHOXAM"))

ggplot(neonics_raw_ohio, aes(x = YEAR, y = KG))+
  geom_point()+
  facet_wrap(~COMPOUND)
#similar to Tyson's. Phew.

````

## Controlling for county area
Metadata:
GCT-PH1 population, Housing Units, Area, and Density:  2010 - State -- County / County Equivalent

NOTE: For information on confidentiality protection, nonsampling error, and definitions, see  http://www.census.gov/prod/cen2010/doc/sf1.pdf.

Available from: http://factfinder2.census.gov/faces/tableservices/jsf/pages/productview.xhtml?src=bkmk 


```{r, results = "hide", cache=FALSE}
ohio_facts<-read.csv("raw_data/DEC_10_SF1_GCTPH1/DEC_10_SF1_GCTPH1.ST05_with_ann.csv")

head(ohio_facts)
names(ohio_facts)
ohio_facts_r <- ohio_facts[ , c(2, 3, 5, 7, 10:14)]
head(ohio_facts_r)
names(ohio_facts_r) <- c("state_fips", "state", "GEOID", "county",
                         "total_area", "total_water", "total_land",
                         "popdens_sqmile", "housedens_sqmile")

#FIPS codes and tidying the double header row!

#now to merge
neonics_raw_ohio$county_id<-str_pad(neonics_raw_ohio$COUNTY_FIPS_CODE, 3, "left", pad = "0")
neonics_raw_ohio$GEOID <- paste(neonics_raw_ohio$STATE_FIPS_CODE, neonics_raw_ohio$county_id, sep = "")

ohio_facts_r <- ohio_facts_r[-c(1,2),]
neonics_facts_ohio <- merge(neonics_raw_ohio, ohio_facts_r)

neonics_facts_ohio$total_land<-as.numeric(neonics_facts_ohio$total_land)
neonics_facts_ohio$kg_density <- neonics_facts_ohio$KG/neonics_facts_ohio$total_land


#not quite the same as Tyson's:
ggplot(neonics_facts_ohio, aes(x = YEAR, y = kg_density)) +
         geom_point()+
         facet_wrap(~COMPOUND)+
         theme_bw()


````

# Data reading done (almost)

````{r, results = "hide", cache=TRUE}
ohio_BBS_birds_summary <- data.frame(ohio_BBS_birds %>%
  group_by(year, GEOID, common_name) %>%
  summarise(mean_abund = mean(route_abundance),
            sd = sd(route_abundance),
            n = n()) %>%
  mutate(sem_abund = sd/sqrt(n)))
```

# Summarising data

````{r}
#which(ohio_BBS_birds_summary$n>1)
#names(ohio_BBS_birds_summary)

ggplot(ohio_BBS_birds_summary, aes(x = year, y = mean_abund))+
  geom_line(aes(colour = GEOID, group = GEOID))+
  facet_wrap(~common_name, scales = "free") +
  theme_bw()+
  theme(legend.position = "bottom")
````

```{r}
names(ohio_all_birds_geo)<-toupper(names(ohio_all_birds_geo))
#need to merge the year
summary(ohio_all_birds_geo$YEAR)
summary(neonics_facts_ohio$YEAR)
sort(unique(ohio_all_birds_geo$YEAR))
sort(unique(neonics_facts_ohio$YEAR))

all_birds_ohio<-merge(ohio_all_birds_geo[ ohio_all_birds_geo$YEAR > 1995 & ohio_all_birds_geo$YEAR < 2010, ], neonics_facts_ohio[neonics_facts_ohio$YEAR > 1995 & neonics_facts_ohio$YEAR < 2010,])


names(ohio_BBS_birds_summary)<-toupper(names(ohio_BBS_birds_summary))
neonics_facts_ohio$GEOID<-as.numeric(neonics_facts_ohio$GEOID)

head(neonics_facts_ohio)
head(ohio_BBS_birds_summary)
ohio_insects <- merge(neonics_facts_ohio[neonics_facts_ohio$YEAR >= 1996 & neonics_facts_ohio$YEAR <=2009,], ohio_BBS_birds_summary[ohio_BBS_birds_summary$YEAR >= 1996 & ohio_BBS_birds_summary$YEAR <=2009,], by = c("YEAR", "GEOID"))

unique(ohio_insects$YEAR)

# without doing it for each county, response looks pretty poisson
ggplot(ohio_BBS_birds_summary, aes(x = MEAN_ABUND))+
  geom_histogram()+
  theme_bw()+
  facet_wrap(~COMMON_NAME, scales = "free")

#more graphs. I had a play. Excluded the two species as very low occurrence.

ggplot(subset(ohio_insects,ohio_insects$YEAR %in% c(2004,2007, 2009) & !ohio_insects$COMMON_NAME %in% c("BLACK TERN", "BLACKPOLL WARBLER")), aes(y = log(MEAN_ABUND), x = log(kg_density)))+
  geom_point(aes(colour = factor(YEAR)))+
  scale_colour_brewer(type = "qual", palette = "Dark2")+
  geom_smooth(aes(colour = factor(YEAR)), method = lm)+
  theme_bw()+
  facet_grid(COMPOUND~COMMON_NAME, scales = "free")+
  coord_fixed()

ggplot(subset(ohio_insects, !ohio_insects$COMMON_NAME %in% c("BLACK TERN", "BLACKPOLL WARBLER")), aes(y = log(MEAN_ABUND), x = YEAR))+
  geom_point(aes(colour = factor(COMMON_NAME)))+
  scale_colour_brewer(type = "qual", palette = "Dark2", guide = FALSE)+
  geom_smooth(aes(colour = factor(COMMON_NAME)), method = lm)+
  theme_bw()+
  facet_grid(COMMON_NAME~COMPOUND, scales = "free")+
  coord_fixed()
````

# First attempt at modelling

```{r}
ohio_insects_wide <- spread(ohio_insects, key = COMPOUND, value = kg_density) 

ohio_insects_wide$GEOID<-as.factor(ohio_insects_wide$GEOID)
ohio_insects_wide$county<-as.factor(ohio_insects_wide$county)
ohio_insects$MEAN_ABUND <- round(ohio_insects$MEAN_ABUND) #poisson requires integers
ohio_insects_wide$MEAN_ABUND <- round(ohio_insects_wide$MEAN_ABUND)

#fails.  I had one working
try(mod1 <- glm(data=ohio_insects_wide, MEAN_ABUND ~ COMMON_NAME:(CLOTHIANIDIN + ACETAMIPRID + IMIDACLOPRID+THIAMETHOXAM) + county, family = poisson))
try(summary(mod1))

### i think i need to go back and put in wide levels of compound and put zero where not recorded######
try(mod1_pred <- predict(mod1, type = "response"))

#works.  But this is not the wide version and only one species, otherwise the interactions get messy.
mod2<- glm(data= ohio_insects[ohio_insects$COMMON_NAME == "CHIMNEY SWIFT",], MEAN_ABUND ~ kg_density:COMPOUND,
           family = "poisson")
summary(mod2)
save.image(file = "BBS_pest.Rdata")
````

```{r exploring landcover}
#setwd("/Users/Liv/Documents/NCEAS_GIT/NCEAS-RENCI_2014/BBS_data")
landcover_all <- read.csv("../Landcover/landcover_all.csv")
head(landcover_all)
head(ohio_insects)

unique(ohio_insects$YEAR)
unique(landcover_all$year)
sort(unique(neonics_facts_ohio$YEAR))
sort(unique(neonics_facts_ohio$YEAR))
sort(unique(ohio_BBS_birds_summary$YEAR))
sort(unique(ohio_BBS_birds$year))
sort(unique(neonics_raw_ohio$YEAR))

class(landcover_all$year)
head(landcover_all)
landcover_sum1 <- data.frame(landcover_all %>%
                               group_by(year, reclass) %>%
                               summarise(
                                 landcover = sum(total)))
landcover_sum2 <-aggregate(total ~ year + reclass, landcover_all,  FUN = sum)

ggplot(landcover_sum2[landcover_sum1$year>1992, ], aes(x = year, y = total)) +
  geom_bar(aes(fill = factor(reclass)), stat = "identity")+
  theme_bw() +
  facet_wrap(~ reclass, scales = "free") +
  scale_fill_colorblind(guide = FALSE)

```

````{r}
head(ohio_BBS_birds)
names(ohio_BBS_birds)

ohio_BBS_birds_short <- ohio_BBS_birds[ ,c(58, 71, 7, 3, 69, 70)] # ignoring the "stop" data primarily
head(ohio_BBS_birds_short)
ohio_BBS_birds_short2 <- spread(ohio_BBS_birds_short, key = common_name, value = route_abundance)
head(ohio_BBS_birds_short2)
names(ohio_BBS_birds_short2)
ohio_BBS_birds_short2[is.na(ohio_BBS_birds_short2)]<-0

head(ohio_BBS_birds_short2)

require(vegan)
names(ohio_BBS_birds_short2)


nmds_neonics <- metaMDS(ohio_BBS_birds_short2[, 5:14])
plot(nmds_neonics)
checking <- str(scores(nmds_neonics, display = "sites"))
birdMDS <- data.frame(NMDS1 = nmds_neonics$points[,1],
                      NMDS2 = nmds_neonics$points[,2])

ohio_BBS_birds_short2[c(which(birdMDS$NMDS1 > 8)), "BLACKPOLL WARBLER"]
#blackpoll warbler is outlier

ordiellipse(nmds_neonics, ohio_BBS_birds_short2$year, col = "forestgreen")
veg<-vegdist(ohio_BBS_birds_short2[, 3:12], method = "bray")

beta_adonis <- adonis(veg ~ year, data = ohio_BBS_birds_short2, permutations = 500, strata = "RouteDataID")
beta_nenonic <- betadisper(veg, ohio_BBS_birds_short2$year)
beta_nenonic

boxplot(beta_nenonic, ylab = "Distance to centroid")

## S3 method for class 'betadisper'
TukeyHSD(beta_nenonic, ordered = TRUE)

head(neonics_raw_ohio)
neonics_raw_ohioj <- spread(neonics_raw_ohio, key = COMPOUND, value = KG) # not working
dim(neonics_raw_ohioj)
dim(neonics_raw_ohio)
dim(ohio_BBS_birds_short2) #merge by rows, exclude the right years.
neonvegbirds <- merge(neonics_raw_ohio, ohio_BBS_birds_short2)

```