---
title: "BBS_neonics_landuse"
author: "Olivia Burge"
date: "21 January 2015"
output: html_document
---

````{r, warning = FALSE, message = FALSE, results='hide'}
require(ggthemes)
require(ggplot2)
require(vegan)
require(dplyr)
require(data.table)
require(tidyr)
````

# Reading in the new bird data
From Tyson's csv. Creates a column for the abundance of birds by route (data is by stop).
````{r}
setwd("/Users/Liv/Documents/NCEAS_GIT/NCEAS-RENCI_2014/BBS_data/")
new_bbs <- read.csv("fifty7_withRTENO.csv")
new_bbs$sum_route_abundance <- rowSums(new_bbs[grep("^Stop[0-9]+", names(new_bbs))])
head(new_bbs)
names(new_bbs)
new_bbs <- new_bbs[ , c(1:7, 58, 59)] #selects everything except the individual stop abundances
````

# Bird trait data
Read in and pare down the bird trait data to the scientific, common names for species and their habitat, nesting, diest, behaviour and conservation status, and 4&6 letter codes.

````{r}
bird_traits <- read.csv("OH_BBS_1966-2013_traits.csv")
bird_traits <- bird_traits[ , c(1,2,3,4,5,6,7,8,9)]
head(bird_traits[2])
names(bird_traits) <- toupper(names(bird_traits))
names(bird_traits) <- gsub("[.]", "_", names(bird_traits))  # using the [] to get the period to be used literally. Or can use "fixed = TRUE"
names(bird_traits)
bird_traits$SCIENTIFIC_NAME <- toupper(bird_traits$SCIENTIFIC_NAME)
head(bird_traits)
````

# Bird code AOU data
BBS data comes with a different kind of AOU code. This chunk tidies up the AOU code dataframe and makes the common and scienitific names compatible with the BBS data. Need to use the common names, scientific names are slightly off in some cases.

```{r}
AOU_codes <- read.csv("raw_data/AOU_codes.csv") ## a cleaned up (ie header removed, split into columns - no substantive changes to any species names or column headers) version of: ftp://ftpext.usgs.gov/pub/er/md/laurel/BBS/DataFiles/SpeciesList.txt

head(AOU_codes)
AOU_codes$common_name <- toupper(AOU_codes$English_Common_Name)
names(AOU_codes)
AOU_codes <- AOU_codes[ , c(2, 10, 6:9)]
names(AOU_codes) <- toupper(names(AOU_codes))
head(AOU_codes)
AOU_codes$SCIENTIFIC_NAME <- paste(AOU_codes$GENUS, AOU_codes$SPECIES, sep = " ")
AOU_codes$SCIENTIFIC_NAME <- toupper(AOU_codes$SCIENTIFIC_NAME)
#testing that all our traits at leat are in the new AOU codes:
bird_traits$SCIENTIFIC_NAME %in% AOU_codes$SCIENTIFIC_NAME # = not all true.
AOU_codes$SCIENTIFIC_NAME %in% bird_traits$SCIENTIFIC_NAME # many many falses

head(bird_traits)
bird_traits$COMMON_NAME %in% AOU_codes$English_Common_Name # = all true!
AOU_codes$English_Common_Name %in% bird_traits$COMMON_NAME # many many falses

unique(new_bbs$AOU) %in% unique(AOU_codes$AOU) # good.

# merge the AOU codes into the ohio data
```

# Merging AOU codes and traits

Putting the trait and AOU codes together - 
```{r}
head(AOU_codes)
head(new_bbs)
AOU_codes$COMMON_NAME <- AOU_codes$English_Common_Name

traits_aou <- inner_join(bird_traits, AOU_codes)

BBS_traits <- inner_join(new_bbs, traits_aou)
names(BBS_traits)
````

# Neonics data by route
Based on Tyson's GIS analysis.
````{r}
neonics_buffers <- read.csv("../Pesticides/pest_buff_overtime.csv")
head(neonics_buffers)

````

# Merge BBS_traits and neonics
```{r}
sort(unique(neonics_buffers$YEAR))
names(BBS_traits) <- toupper(names(BBS_traits))
sort(unique(BBS_traits$YEAR))
````
Therefore, we need to select the period 1995 - 2011 for merging.

```{r}
BBS_neonics <- inner_join(BBS_traits[ BBS_traits$YEAR < 2012, ], neonics_buffers[ neonics_buffers$YEAR > 1994, ]) # joins by RTENO and YEAR which works. 
````

# Landuse data by route
Based on Tyson's GIS analysis. Notes for the csv: `reclass` is Anderson Level I land cover: (0=nodata, 1=water, 2=developed, 3=barren, 4=forest, 5=grassland/shrub, 6=agriculture, 7=wetlands).
`total` is the number of pixels (multiply by 0.009 to get square km).
````{r}
landuse_buffers <- read.csv("../Pesticides/LC_buffers_overtime.csv")
head(landuse_buffers)

reclass_table <- data.frame("reclass" = 0:7, "land_cover" = c("nodata", "water", "developed", "barren", "forest", "grassland_shrub", "agriculture", "wetlands"))

landuse_buffers <- merge(landuse_buffers, reclass_table, by = "reclass")
class(landuse_buffers$land_cover)

head(landuse_buffers)
tail(landuse_buffers)

````
````{r}
names(landuse_buffers)
landuse_buffers2 <- landuse_buffers[c(2, 3, 4, 5, 7, 8)]
names(landuse_buffers2)
class(landuse_buffers2$total_pix)
str(landuse_buffers2)
head(landuse_buffers2)

landuse_buffers_wide <- spread(landuse_buffers2, land_cover, total_pix)
tail(landuse_buffers_wide, 50)

names(landuse_buffers_wide)
````
# Merging BBS_neonics with landuse
```{r}
sort(unique(landuse_buffers$YEAR)) # 1992 - 2011

BBS_ln <- inner_join(BBS_neonics, landuse_buffers_wide[ landuse_buffers_wide$YEAR > 1994, ])
# joins by: c("YEAR", "RTENO", "BBS_route", "buffer")

tail(BBS_ln)
```

# Maybe a baby model?

```{r}
names(BBS_ln)
unique(BBS_ln$buffer)

BBS_ln$agriculture_scaled <- scale(BBS_ln$agriculture)
BBS_ln$high_kg_buff_scaled <- scale(BBS_ln$high_kg_buff)
head(BBS_ln)
test_mod2 <- glmer(data = BBS_ln[ BBS_ln$buffer == 200, ],
                   family = poisson,
                   formula= SUM_ROUTE_ABUNDANCE ~ agriculture_scaled + high_kg_buff_scaled:COMPOUND + DIET + (1|BBS_route) + (1|YEAR))
summary(test_mod2)
require(coefplot2)
coefplot2(test_mod2)


require(dismo)
BBS_200 <- BBS_ln[ BBS_ln$buffer == 200, ]
names(BBS_200)
class(BBS_200$SPECIES)
gbm_neonic <- gbm.step(data = BBS_200, 
                     gbm.x=c(1, 3, 15, 16, 17, 18, 23, 25, 26, 29:35), 
                     gbm.y = 11, family = "poisson",
                     tree.complexity = 14) #, 
                      #step.size = 20, n.trees = 50,
                     #learning.rate = 0.001, bag.fraction = 0.75)
summary(gbm_neonic)

glmm_neonic <- glmer(data = BBS_200,
                     formula = SUM_ROUTE_ABUNDANCE ~ high_kg_buff:COMPOUND + 
                       high_kg_buff:forest:COMPOUND + 
                       high_kg_buff:agriculture:COMPOUND + 
                       forest + agriculture +
                       (1|GENUS) + (1|YEAR),
                     family = "poisson")

summary(glmm_neonic)
coefplot2(glmm_neonic)

neonic_int<-gbm.interactions(gbm_neonic)
neonic_int$interactions

gbm.perspec(gbm_neonic, 1,3)
gbm.perspec(gbm_neonic, 2,3)
gbm.perspec(gbm_neonic, 4,8)
gbm.plot(gbm_neonic, 3,4)

save.image(file = "big_gbm.Rdata")
load(file = "big_gbm.Rdata")
#saveRDS(gbm_neonic, file = "big_serialised_gbm.rds")

```

# Some exploration! 
```{r}
head(landuse_buffers)

ggplot(landuse_buffers[landuse_buffers$land_cover %in% c("forest", "agriculture", "developed", "grassland_shrub") & landuse_buffers$buffer == 400, ], aes(x = YEAR, y = total_pix)) +
  geom_smooth(aes(colour = land_cover)) +
#  geom_density2d(aes(colour = land_cover)) +
  geom_point(aes(colour = land_cover), alpha = 0.3, position = position_jitter(width = 0.3))+
  theme_bw()+
  scale_colour_colorblind("Land Cover", labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub"))+
  guides(colour=guide_legend(override.aes=list(fill=NA)))+
  theme(legend.position = "bottom", legend.key = element_blank(), legend.background = element_rect(colour = "grey"))

ggplot(landuse_buffers[landuse_buffers$land_cover %in% c("forest", "agriculture", "developed", "grassland_shrub") & landuse_buffers$buffer == 400, ], aes(x = YEAR, y = total_pix)) +
  geom_smooth(aes(colour = land_cover)) +
  geom_density2d(aes(colour = land_cover)) +
  #geom_point(aes(colour = land_cover), alpha = 0.3, position = position_jitter(width = 0.3))+
  theme_bw()+
  scale_colour_colorblind("Land Cover", labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub"))+
  guides(colour = FALSE) +
  facet_wrap(~ land_cover, scales = "free")

ggplot(landuse_buffers[landuse_buffers$land_cover %in% c("forest", "agriculture", "developed", "grassland_shrub") & landuse_buffers$buffer == 400 & landuse_buffers$YEAR %in% c(1995, 2000, 2005, 2010), ], aes(x = land_cover, y = total_pix)) +
 geom_violin(aes(fill = land_cover))+
  theme_bw()+
  scale_fill_colorblind("Land Cover", labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
  facet_wrap(~YEAR)


ggplot(landuse_buffers[landuse_buffers$land_cover %in% c("forest", "agriculture", "developed", "grassland_shrub") & landuse_buffers$buffer == 400 & landuse_buffers$YEAR %in% c(1995, 2000, 2005, 2010), ], aes(x = factor(YEAR), y = total_pix)) +
 geom_violin(aes(fill = land_cover))+
  theme_bw()+
  scale_fill_colorblind("Land Cover", labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
  guides(fill = FALSE)+
  facet_wrap(~land_cover, scales = "free")


summary_landuse <- landuse_buffers[landuse_buffers$land_cover %in% c("forest", "agriculture", "developed", "grassland_shrub"), ] %>% group_by(YEAR, land_cover, buffer) %>% filter(buffer == 400) %>% summarise(total_pix = sum(total_pix))

ggplot(summary_landuse, aes(x = YEAR, y = total_pix)) +
  geom_area(aes(fill = land_cover, colour = land_cover), position = "fill")+
  theme_bw()+
  scale_fill_manual("Land Cover", values = c("sienna4", "gray20", "forestgreen", "orange"), labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
  scale_colour_manual("Land Cover", values = c("sienna4", "gray20", "forestgreen", "orange"), labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
  theme(legend.position = "bottom", legend.background = element_rect(colour = "grey"))

unique(landuse_buffers$buffer)
summary_landuse3 <- landuse_buffers[landuse_buffers$land_cover %in% c("forest", "agriculture", "developed", "grassland_shrub"), ] %>% group_by(YEAR, land_cover, buffer) %>% summarise(total_pix = sum(total_pix))

ggplot(summary_landuse3, aes(x = YEAR, y = total_pix)) +
  geom_area(aes(fill = land_cover, colour = land_cover), position = "fill")+
  theme_bw()+
  scale_fill_manual("Land Cover", values = c("sienna4", "gray20", "forestgreen", "orange"), labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
  scale_colour_manual("Land Cover", values = c("sienna4", "gray20", "forestgreen", "orange"), labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
  theme(legend.position = "bottom", legend.background = element_rect(colour = "grey")) +
  facet_wrap(~buffer)


# summary_landuse2 <- landuse_buffers%>% group_by(YEAR, land_cover, buffer) %>% filter(buffer == 400) %>% summarise(total_pix = sum(total_pix))
# 
# ggplot(summary_landuse2, aes(x = YEAR, y = total_pix)) +
#   geom_area(aes(fill = land_cover, colour = land_cover), position = "fill")+
#   theme_bw()+
#   #scale_fill_manual("Land Cover", values = c("sienna4", "gray20", "forestgreen", "orange"), labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
#   #scale_colour_manual("Land Cover", values = c("sienna4", "gray20", "forestgreen", "orange"), labels = c("Agriculture", "Developed", "Forest", "Grass/Shrub")) +
#   theme(legend.position = "bottom", legend.background = element_rect(colour = "grey"))

````

# Exploring the neonics

```{r}
head(neonics_buffers)

head(neonics_buffers[is.na(neonics_buffers$high_kg_buff) | is.na(neonics_buffers$ag_pix_buff) | is.na(neonics_buffers$low_kg_buff), ])
head(neonics_buffers[is.na(neonics_buffers$high_kg_buff) | is.na(neonics_buffers$ag_pix_buff) , ])
dim(neonics_buffers[is.na(neonics_buffers$high_kg_buff) | is.na(neonics_buffers$ag_pix_buff) | is.na(neonics_buffers$low_kg_buff), ])
dim(neonics_buffers[is.na(neonics_buffers$high_kg_buff) | is.na(neonics_buffers$ag_pix_buff), ])

unique(neonics_buffers$YEAR)
class(neonics_buffers$high_kg_buff)
noyear_neonics_buffers <- neonics_buffers %>% 
  group_by(BBS_route, buffer) %>% 
  summarise(high_pest = sum(high_kg_buff), agpix = sum(ag_pix_buff))
head(noyear_neonics_buffers)
# spps <- spps[!is.na(spps$NMDS1) & !is.na(spps$NMDS2),]
noyear_neonics_buffers <- noyear_neonics_buffers[!is.na(noyear_neonics_buffers$high_pest) & !is.na(noyear_neonics_buffers$agpix), ]

ggplot(noyear_neonics_buffers, aes(y = as.numeric(high_pest/agpix), x = as.numeric(buffer))) +
  geom_line(aes(group = BBS_route)) +
  geom_smooth()+
  theme_bw()

ggplot(noyear_neonics_buffers, aes(y = high_pest, x = buffer)) +
  geom_line(aes(group = BBS_route)) +
  geom_smooth()+
  theme_bw()

ggplot(noyear_neonics_buffers, aes(y = agpix, x = buffer)) +
  geom_line(aes(group = BBS_route)) +
  geom_smooth()+
  theme_bw()


# now to see if intensity has increased over tine
names(neonics_buffers)
year_neonics_buffers <- neonics_buffers %>%
  group_by(BBS_route, YEAR, COMPOUND, buffer) %>%
  summarise(intensity = high_kg_buff/ag_pix_buff) 

names(year_neonics_buffers)

ggplot(year_neonics_buffers, aes(y = intensity, x = YEAR)) +
  geom_line(aes(group = BBS_route)) +
  facet_grid(buffer~ COMPOUND, scale = "free") +
  theme_bw()

year_neonics_buffers2 <- neonics_buffers %>%
  group_by(BBS_route, YEAR, COMPOUND) %>%
  summarise(intensity = mean(high_kg_buff/ag_pix_buff)) 

ggplot(year_neonics_buffers2, aes(y = intensity, x = YEAR)) + 
  geom_line(aes(group = BBS_route)) +
  facet_wrap(~ COMPOUND, scale = "free") +
  theme_bw()

ggplot(year_neonics_buffers2, aes(y = intensity, x = YEAR)) + 
  geom_line(aes(group = BBS_route)) +
  facet_wrap(~ COMPOUND) +
  theme_bw()

```

# Using the big dataset: BBS_ln
```{r}
head(BBS_ln)
names(BBS_ln) <- toupper(names(BBS_ln))
sort(unique(BBS_ln$DIET))
BBS_ln$DIET <- relevel(BBS_ln$DIET, ref = "insects")
BBS_ln$YEAR_SCALED <- scale(BBS_ln$YEAR)
BBS_ln$FOREST_SCALED <- scale(BBS_ln$FOREST)
BBS_ln$AGRICULTURE_SCALED <- scale(BBS_ln$AGRICULTURE)
BBS_ln$FOREST_PROP <- BBS_ln$FOREST/BBS_ln$AGRICULTURE
BBS_ln$HIGH_KG_BUFF <- scale(BBS_ln$HIGH_KG_BUFF)
BBS_ln$DIET_VULN <- ifelse(BBS_ln$DIET %in% c("insects", "seeds", 'nectar', "birds"), "diet_vuln", "diet_invuln")
BBS_ln$DIET_VULN <- factor(BBS_ln$DIET_VULN)

BBS_ln$DIET_INSECTS <- ifelse(BBS_ln$DIET == "insects", "diet_insects", "diet_not_insects")
BBS_ln$DIET_INSECTS <- factor(BBS_ln$DIET_INSECTS)

with(BBS_ln, cor(FOREST, AGRICULTURE))
test_mod <- glm(SUM_ROUTE_ABUNDANCE ~ YEAR_SCALED + 
                  AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
                  AGRICULTURE_SCALED +
                  DIET_VULN:YEAR_SCALED + 
                  DIET_VULN:HIGH_KG_BUFF_SCALED:YEAR_SCALED:AGRICULTURE_SCALED + 
                  HIGH_KG_BUFF_SCALED:COMPOUND + 
                  HIGH_KG_BUFF_SCALED, 
                data = BBS_ln[BBS_ln$COMPOUND %in% c("ATRAZINE", "GLYPHOSATE"),], 
                family = "poisson")
summary(test_mod)
coefplot2(test_mod)
#require(effects)
plot(allEffects(test_mod))

test_mod2 <- glm(SUM_ROUTE_ABUNDANCE ~ YEAR_SCALED + 
                  AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
                  AGRICULTURE_SCALED +
                  DIET_INSECTS:YEAR_SCALED + 
                  DIET_INSECTS:HIGH_KG_BUFF_SCALED:YEAR_SCALED:AGRICULTURE_SCALED + 
                  HIGH_KG_BUFF_SCALED:COMPOUND + 
                  HIGH_KG_BUFF_SCALED, 
                data = BBS_ln[BBS_ln$COMPOUND %in% c("ATRAZINE", "GLYPHOSATE"),], 
                family = "poisson")
summary(test_mod2)
coefplot2(test_mod2)


glmer_mod_3 <- glmer(SUM_ROUTE_ABUNDANCE ~ YEAR_SCALED + 
                  AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
                  AGRICULTURE_SCALED +
                  DIET_INSECTS:YEAR_SCALED + 
                  DIET_INSECTS:HIGH_KG_BUFF_SCALED:YEAR_SCALED:AGRICULTURE_SCALED + 
                  HIGH_KG_BUFF_SCALED:COMPOUND + 
                  HIGH_KG_BUFF_SCALED +
                    (1|ROUTE), 
                data = BBS_ln[BBS_ln$COMPOUND %in% c("ATRAZINE", "GLYPHOSATE"),], 
                family = "poisson")
coefplot2(glmer_mod_3)
coefplot2(list(glmer_mod_3, test_mod2))

glmer_mod_4 <- glmer(SUM_ROUTE_ABUNDANCE ~ YEAR_SCALED + 
                  AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
                  AGRICULTURE_SCALED +
                  DIET_INSECTS:YEAR_SCALED + 
                    DIET_INSECTS +
                  DIET_INSECTS:HIGH_KG_BUFF_SCALED:YEAR_SCALED:AGRICULTURE_SCALED + 
                  HIGH_KG_BUFF_SCALED:COMPOUND + 
                  HIGH_KG_BUFF_SCALED +
                    (1|ROUTE), 
                data = BBS_ln[BBS_ln$COMPOUND %in% c("ATRAZINE", "GLYPHOSATE"),], 
                family = "poisson")
summary(glmer_mod_4)
coefplot2(glmer_mod_4)
coefplot2(list(glmer_mod_3, test_mod2, glmer_mod_4))

glmer_mod_5 <- glmer(SUM_ROUTE_ABUNDANCE ~ 
                       AGRICULTURE_SCALED:DIET_INSECTS + 
                       AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
                       FOREST_SCALED:DIET_INSECTS +
                       DIET_INSECTS:HIGH_KG_BUFF_SCALED:AGRICULTURE_SCALED + 
                       HIGH_KG_BUFF_SCALED:COMPOUND:DIET_INSECTS + 
                       HIGH_KG_BUFF_SCALED:DIET_INSECTS +
                       (ROUTE|YEAR), 
                data = BBS_ln[BBS_ln$COMPOUND %in% c("ATRAZINE", "GLYPHOSATE"),], 
                family = "poisson")
summary(glmer_mod_5)
coefplot2(glmer_mod_5)
coefplot2(list(glmer_mod_3, test_mod2, glmer_mod_4, glmer_mod_5))
```

# Aggregating based on diet and then modelling

```{r}
head(BBS_ln)
names(BBS_ln)
BBS_ln$HIGH_KG_BUFF <- as.numeric(BBS_ln$HIGH_KG_BUFF)
BBS_ln$AGRICULTURE_SCALED <- as.numeric(BBS_ln$AGRICULTURE_SCALED)
BBS_ln$HIGH_KG_BUFF_SCALED <- as.numeric(BBS_ln$HIGH_KG_BUFF_SCALED)
BBS_ln$YEAR_SCALED <- as.numeric(BBS_ln$YEAR_SCALED)
BBS_ln$FOREST_SCALED <- as.numeric(BBS_ln$FOREST_SCALED)
BBS_ln_diet <- data.frame(BBS_ln %>% group_by(YEAR, RTENO, BBS_ROUTE, BUFFER, ROUTE, COMPOUND, HIGH_KG_BUFF, AGRICULTURE, FOREST, DEVELOPED, AGRICULTURE_SCALED, YEAR_SCALED, FOREST_SCALED, HIGH_KG_BUFF_SCALED, FOREST_PROP, DIET) %>%
  summarise(SUM_ROUTE_ABUNDANCE_DIET = sum(SUM_ROUTE_ABUNDANCE)))
head(BBS_ln_diet)

````

```{r}
class(BBS_ln_diet$ROUTE)
BBS_ln_diet$YEAR_fac <- as.factor(BBS_ln_diet$YEAR)
BBS_ln_diet$ROUTE <- as.factor(BBS_ln_diet$ROUTE)


range(na.omit(BBS_ln_diet$HIGH_KG_BUFF_SCALED))
range(na.omit(BBS_ln_diet$AGRICULTURE_SCALED))
range(na.omit(BBS_ln_diet$FOREST_SCALED))
range(na.omit(BBS_ln_diet$FOREST_SCALED))
class(BBS_ln_diet$DIET)
class(BBS_ln_diet$ROUTE)
range(BBS_ln_diet$SUM_ROUTE_ABUNDANCE_DIET)
ggplot(BBS_ln_diet, aes(x = ROUTE, y = SUM_ROUTE_ABUNDANCE_DIET)) +
  geom_point() +
  geom_boxplot() + theme_bw() +
  facet_wrap(~ YEAR)

BBS_ln_diet$sqrt_response <- round(sqrt(BBS_ln_diet$SUM_ROUTE_ABUNDANCE_DIET), 0)
range(BBS_ln_diet$sqrt_response)

BBS_ln_dietcc <- BBS_ln_diet[complete.cases(BBS_ln_diet),]
str(BBS_ln_dietcc)
glmer_mod_6 <- glmer(na.action = na.omit,
                     sqrt_response ~ 
                       AGRICULTURE_SCALED:DIET+ 
                       AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
                       FOREST_SCALED:DIET +
                       DIET:HIGH_KG_BUFF_SCALED:AGRICULTURE_SCALED + 
                       HIGH_KG_BUFF_SCALED:COMPOUND:DIET + 
                       HIGH_KG_BUFF_SCALED:DIET +
                       (1|YEAR_fac) + (1|ROUTE), 
                     data = BBS_ln_dietcc[BBS_ln_dietcc$COMPOUND %in% c("ATRAZINE", 
                                                                    "GLYPHOSATE")
                                        & BBS_ln_dietcc$YEAR > 1996
                                        & BBS_ln_dietcc$DIET != "",], 
                     family = "poisson", 
                     verbose = TRUE)
sum_glmer_mod_6<-summary(glmer_mod_6)
cf2_glmer_mod_6 <- coefplot2(glmer_mod_6)
# save(glmer_mod_6, sum_glmer_mod_6, cf2_glmer_mod_6, file = "glmer_mod_6.Rdata")
#load( "glmer_mod_6.Rdata")
#save.image("predicting_glmers.Rdata")
diet_levels <- levels(BBS_ln_dietcc$DIET)[-2]
ep <- expand.grid(AGRICULTURE_SCALED = seq(from = min(BBS_ln_dietcc$AGRICULTURE_SCALED), 
                                    to = max(BBS_ln_dietcc$AGRICULTURE_SCALED),
                                    by = 0.4),
                  FOREST_SCALED = mean(BBS_ln_dietcc$FOREST_SCALED),
                  DIET = diet_levels, 
                  HIGH_KG_BUFF_SCALED =seq(from = min(BBS_ln_dietcc$HIGH_KG_BUFF_SCALED), 
                                    to = max(BBS_ln_dietcc$HIGH_KG_BUFF_SCALED),
                                    by = 2),
                  COMPOUND = c("GLYPHOSATE", "ATRAZINE")
                    )

mm <- model.matrix(delete.response(terms(glmer_mod_6)), ep)

ep <- data.frame(ep, response = mm %*% fixef(glmer_mod_6))
ep <- with(ep, data.frame(ep, transformed.response = exp(response)))

pvar1 <- diag(mm %*% tcrossprod(vcov(glmer_mod_6), mm))
tvar1 <- pvar1 + VarCorr(glmer_mod_6)$YEAR + VarCorr(glmer_mod_6)$ROUTE
ep <- data.frame(ep, pse=sqrt(pvar1), tse=sqrt(tvar1))
ep <- with(ep,
                data.frame(ep,
                           plo=exp(response-1.96*pse),
                           phi=exp(response+1.96*pse),
                           tlo=exp(response-1.96*tse),
                           thi=exp(response+1.96*tse)))
head(ep)

names(ep) <- tolower(names(ep))
names(ep) 
ep1 <- ep[ep$diet %in% c("insects", "omnivore"), ]
require(akima)
ep2 <- with(ep1[ep1$compound == "GLYPHOSATE" & ep1$diet == "insects",], interp(agriculture_scaled, high_kg_buff_scaled, transformed.response))
ep3 <- with(ep1[ep1$compound == "GLYPHOSATE" & ep1$diet == "omnivore",], interp(agriculture_scaled, high_kg_buff_scaled, transformed.response))

ep4 <- with(ep1[ep1$compound == "ATRAZINE" & ep1$diet == "insects",], interp(agriculture_scaled, high_kg_buff_scaled, transformed.response))
ep5 <- with(ep1[ep1$compound == "ATRAZINE" & ep1$diet == "omnivore",], interp(agriculture_scaled, high_kg_buff_scaled, transformed.response))


with(ep2, persp(x, y, z, theta = 330, phi = 20, 
                xlab = "Agriculture",
                ylab = "Pesticide application rate", 
                zlab = "Bird response",
                col = "royalblue",
                #nticks = 5, 
                #ticktype = "detailed",
                shade = 0.75))

with(ep3, persp(x, y, z, theta = 55, phi = 30, 
                xlab = "Agriculture",
                ylab = "Pesticide application rate", 
                zlab = "Bird response",
                col = "cornflowerblue",
                #nticks = 5, 
                #ticktype = "detailed",
                shade = 0.75))

with(ep4, persp(x, y, z, theta = 280, phi = 30, 
                xlab = "Agriculture",
                ylab = "Pesticide application rate", 
                zlab = "Bird response",
                col = "orange",
                #nticks = 5, 
                #ticktype = "detailed",
                shade = 0.75))

with(ep5, persp(x, y, z, theta = 55, phi = 30, 
                xlab = "Agriculture",
                ylab = "Pesticide application rate", 
                zlab = "Bird response",
                col = "violetred",
                #nticks = 5, 
                #ticktype = "detailed",
                shade = 0.2))


```

Earlier glm model which you can run if interested, heavily pseudo-replicated however.
```{r}
glm_mod_6 <- glm(na.action = na.omit,
                 sqrt_response ~ 
                   DIET+
                   HIGH_KG_BUFF_SCALED +
                   AGRICULTURE_SCALED:DIET+ 
                   AGRICULTURE_SCALED:HIGH_KG_BUFF_SCALED + 
                   DIET:HIGH_KG_BUFF_SCALED:AGRICULTURE_SCALED + 
                   HIGH_KG_BUFF_SCALED:COMPOUND:DIET + 
                   HIGH_KG_BUFF_SCALED:DIET, 
                 data = BBS_ln_dietcc[BBS_ln_dietcc$COMPOUND %in% c("ATRAZINE", 
                                                                    "GLYPHOSATE")
                                      & BBS_ln_dietcc$YEAR > 1996
                                      & BBS_ln_dietcc$DIET != "",], 
                 family = "poisson")
summary(glm_mod_6)
coefplot2(glm_mod_6)

```