---
title: "County landuse change"
author: "Olivia Burge"
date: "1 July 2015"
output: html_document
---

# Data read-in

`LC_county_overtime.csv` has estimated landcover for each OH county and year.

columns descriptions for these 3:
"reclass" is Anderson Level I land cover: (0=nodata, 1=water, 2=developed, 3=barren, 4=forest, 5=grassland/shrub, 6=agriculture, 7=wetlands)
"buffer" is the length of the radius in meters used around transects or sites
"total_pix" is the number of Landsat 30x30m pixels in a grouping
"km2" is derived by multiplying pixels by 0.009 to get square kilometers.
"RTENO" is BBS unique ID for transects
```{r}
setwd("/Users/Liv/Documents/NCEAS_GIT/NCEAS-RENCI_2014/Landcover/")
require(ggplot2); theme_set(theme_bw())
require(gridExtra)
require(tidyr)
require(dplyr)
require(maps)

LC_county <- read.csv("OH_LC_time.csv")
head(LC_county)
geoids <- LC_county %>%
  group_by(GEOID) %>%
  select(GEOID, NAME) %>%
  slice(1)
tail(geoids)
geoids$NAME <- tolower(geoids$NAME)

map_data_ohio <- map_data("county", "ohio")
head(map_data_ohio)
head(geoids)
base_map <- merge(geoids, map_data_ohio, by.x = "NAME", by.y = "subregion", all.y = FALSE)

LC_county_overtime <- read.csv("LC_county_overtime.csv", row.names = 1)
LC_county_overtime$landclass <- ifelse(LC_county_overtime$reclass == 1, "water",
                                       ifelse(LC_county_overtime$reclass == 2, "developed",
                                              ifelse(LC_county_overtime$reclass == 3, "barren",
                                                     ifelse(LC_county_overtime$reclass==4, "forest",
                                                            ifelse(LC_county_overtime$reclass == 5, "grass_shrub", ifelse(LC_county_overtime$reclass == 6, "agriculture", ifelse(LC_county_overtime$reclass == 7, "wetlands", "nodata")))))))
head(LC_county_overtime)
LC_county_overtime_wide <- LC_county_overtime %>% 
  select(GEOID, YEAR, pix, landclass) %>% 
  spread(landclass, pix) 
names(LC_county_overtime_wide)
LC_county_overtime_wide$total <- rowSums(LC_county_overtime_wide[3:9])

geoids2 <- data.frame(LC_county_overtime %>% group_by(GEOID) %>% slice(1))


map_data_LC <- merge(LC_county_overtime_wide,geoids)
map_data_LC$NAME <- tolower(map_data_LC$NAME)


```

```{r}
require(maps)

map_data_merged <- merge(map_data_LC, map_data_ohio, by.x = "NAME", by.y = "subregion")
names(map_data_merged)

map_data_merged_summary_forest<- map_data_merged %>%
  filter(YEAR %in% c(1992, 1997, 2002, 2012)) %>%
  select(NAME, GEOID, YEAR, forest, lat, long, group, order) %>%
  spread(YEAR, forest) %>%
  mutate(change_9297 =  `1997` / `1992`,
         change_9202 =  `2002` / `1992`,
         change_9212 = `2012` / `1992`) %>%
  select(NAME, GEOID, lat, long, group, order, change_9297,change_9202, change_9212) %>%
  gather(key = year, value = change_from_92, change_9297:change_9212)
map_data_merged_summary_agri<- map_data_merged %>%
  filter(YEAR %in% c(1992, 1997, 2002, 2012)) %>%
  select(NAME, GEOID, YEAR, agriculture, lat, long, group, order) %>%
  spread(YEAR, agriculture) %>%
  mutate(change_9297 =  `1997` / `1992`,
         change_9202 =  `2002` / `1992`,
         change_9212 = `2012` / `1992`) %>%
  select(NAME, GEOID, lat, long, group, order, change_9297,change_9202, change_9212) %>%
  gather(key = year, value = change_from_92, change_9297:change_9212)
  
  
head(map_data_merged)

forest_change <- ggplot(map_data_merged_summary_forest, aes(x = long, y = lat, group = NAME)) +
  facet_wrap(~year) +
  geom_polygon(aes(fill = change_from_92)) +
  scale_fill_gradient2(low = "red", mid="grey", high = "forestgreen", midpoint = 1)+
  coord_map()
head(map_data_merged_summary_agri)
agri_change <- ggplot(map_data_merged_summary_agri, aes(x = long, y = lat, group = NAME)) +
  facet_wrap(~year) +
  geom_polygon(aes(fill = change_from_92)) +
  scale_fill_gradient2(low = "blue", mid = "grey", midpoint = 1, high = "orange")+
  coord_map()

agri_forest <- arrangeGrob(forest_change, agri_change, nrow = 2)
agri_forest

```

```{r, echo=FALSE}
pest_intensity <- read.csv("../Pesticides/estimated_pesticides_OHcounty_MOA.csv")

pest_intensity2 <- pest_intensity %>% select(GEOID, YEAR, MOAuse, high_kg_km2ag) %>% ungroup() %>%
  filter(YEAR%in% c(1992, 1997, 2002, 2012)) %>%
  group_by(GEOID, YEAR, MOAuse)%>%
  summarise(mean_application = mean(high_kg_km2ag))


base_map_short <- base_map %>% select(GEOID, NAME, long, lat, group)
head(base_map_short)
head(pest_intensity2)
pest_intensity3 <-   pest_intensity2 %>% ungroup() %>% filter (pest_intensity2$GEOID %in% base_map$GEOID & MOAuse == "her_K")%>%
  select(GEOID, YEAR, MOAuse, mean_application) %>%
                                left_join(., base_map_short)

head(pest_intensity3)
map_data_merged_summary_pest <- pest_intensity3 %>%
    select(NAME, GEOID, YEAR, mean_application, lat, long, group) %>%
  ungroup() %>%
  filter(YEAR %in% c(1992, 1997, 2002, 2012)) %>%
    group_by(GEOID, YEAR, long, lat) %>%
  slice(1) %>%
    ungroup() %>%
  spread(YEAR, mean_application) %>%
  mutate(change_9297 =  `1997` / `1992`,
         change_9202 =  `2002` / `1992`,
         change_9212 = `2012` / `1992`) %>%
  select(NAME, GEOID, lat, long, group, order, change_9297,change_9202, change_9212) %>%
  gather(key = year, value = change_from_92, change_9297:change_9212)

head(base_map)
county_mouse <- ggplot(base_map, aes(x = long, y = lat, group = GEOID)) +
  geom_polygon(aes(mean_application))+
  #facet_wrap(~YEAR) +
  coord_map()
county_mouse
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
