---
title: "Pesticides"
author: "Tyson Wepprich"
date: "Monday, July 28, 2014"
output: pdf_document
---
Goal:
County-level pesticide use data in 2007 from NC, OH, MI, CA
(Are we worried about which chemicals in particular? Or maybe just top 10 most used? PCA?)

Data source:
Estimated Annual Agricultural Pesticide Use for Counties of the Conterminous United States, 1992-2009
http://pubs.usgs.gov/ds/752/

Thelin, G.P., and Stone, W.W., 2013, Estimation of annual agricultural pesticide use for counties of the conterminous United States, 1992-2009: U.S. Geological Survey Scientific Investigations Report 2013-5009, 54 p.


Census 2010 County Data from https://www.census.gov/geo/maps-data/data/gazetteer2010.html
Column  Label	Description
Column 1	USPS 	United States Postal Service State Abbreviation
Column 2	GEOID 	Geographic Identifier - fully concatenated geographic code (State FIPS and County FIPS)
Column 3 	ANSICODE 	American National Standards Institute code
Column 4	NAME 	Name
Column 5 	POP10 	2010 Census population count.
Column 6 	HU10 	2010 Census housing unit count.
Column 7	ALAND 	Land Area (square meters) - Created for statistical purposes only.
Column 8	AWATER 	Water Area (square meters) - Created for statistical purposes only.
Column 9	ALAND_SQMI 	Land Area (square miles) - Created for statistical purposes only.
Column 10	AWATER_SQMI 	Water Area (square miles) - Created for statistical purposes only.
Column 11	INTPTLAT 	Latitude (decimal degrees) First character is blank or "-" denoting North or South latitude respectively.
Column 12 	INTPTLONG 	Longitude (decimal degrees) First character is blank or "-" denoting East or West longitude respectively.

```{r combine text files, echo = FALSE}
setwd("C:/Users/Tyson/Desktop/Pesticides")

filenames <- list.files()
EPA.high <- do.call("rbind", lapply(filenames[grep("high", filenames)], read.table, header = TRUE, sep = "\t"))
EPA.low <- do.call("rbind", lapply(filenames[grep("low", filenames)], read.table, header = TRUE, sep = "\t"))

pest.states.low <- EPA.low[which(EPA.low$STATE_FIPS_CODE %in% c(6, 26, 37, 39)), ]
pest.states.high <- EPA.high[which(EPA.high$STATE_FIPS_CODE %in% c(6, 26, 37, 39)), ]

fips <- read.csv("fips_codes.txt", header = TRUE)
names(fips) <- c("State", "STATE_FIPS_CODE", "COUNTY_FIPS_CODE", "County.Name", "ANSI.class")
fips <- fips[which(fips$STATE_FIPS_CODE %in% c(6, 26, 37, 39)), ]
#create new column for ID to match census data
fips$GEOID <- paste(fips$STATE_FIPS_CODE, formatC(fips$COUNTY_FIPS_CODE, width = 3, format = "d", flag = "0"), sep = "")

county.data.low <- merge(pest.states.low, fips)
county.data.high <- merge(pest.states.high, fips)

pest.2007.high <- county.data.high[which(county.data.high$YEAR == 2007), ]
pest.2007.low <- county.data.low[which(county.data.low$YEAR == 2007), ]

#Bring in county data from 2010 US Census (from factfinder online database)
census <- read.table("Gaz_counties_national.txt", header = TRUE, sep = "\t")

pest.2007.h <- merge(pest.2007.high, census)
pest.2007.l <- merge(pest.2007.low, census)
# 
# total.kg <- tapply(pest.2007.high$KG, pest.2007.high$COMPOUND, sum)
# hist(total.kg[total.kg>10000], breaks = 1000)

write.csv(pest.2007.h, file = "pest.2007.high.csv", row.names = FALSE)
write.csv(pest.2007.l, file = "pest.2007.low.csv", row.names = FALSE)


```
Right tail skew on pesticide distribution, sulfur is #1.
Try PCA by county?


```{r distribution of use}
setwd("C:/Users/Tyson/Desktop/Pesticides")
library(tidyr)
library(dplyr)
library(reshape2)
data <- read.csv("pest.2007.high.csv", header = TRUE)

#These lines find the top 10 pesticides by use in each county
test <- data %>% mutate(County.density = KG / ALAND) %>% group_by(State) %>% mutate(Num.Counties = length(unique(GEOID)))
test2 <- test %>% group_by(GEOID) %>% mutate(County.rank = rank(-County.density))
test3 <- test2 %>% filter(County.rank <=10) %>% data.frame()
TopCounty <- droplevels.data.frame(test3)

a <- TopCounty %>% group_by(COMPOUND, State) %>% summarise(prop.counties = n()/Num.Counties[1]) %>% data.frame()
b <- a %>% arrange(State, prop.counties)
chems <- unique(b$COMPOUND[b$prop.counties >= .5])

chem.data <- TopCounty[which(TopCounty$COMPOUND %in% chems), ]

# Top10State <- data %>% group_by(State, COMPOUND) %>% summarise(State.Total = sum(KG)) %>% group_by(State) %>% mutate(State.rank = rank(-State.Total)) %>% filter(State.rank <= 10)
# 
# Top5NC.County <- data %>% filter(State == "NC") %>% group_by(GEOID) %>% mutate(County.rank = rank(-KG)) %>% filter(County.rank <=5)
# Top5 <- Top5NC.County[, c("COMPOUND", "KG", "County.Name", "County.rank")]


pca.data <- chem.data[, c("GEOID", "State", "County.density", "COMPOUND")]
test <- dcast(pca.data, GEOID + State ~ COMPOUND, mean, value.var = "County.density", fill = 0)

#test <- chem.data %>% group_by(GEOID) %>% summarise(COMPOUND, County.density, fill = 0) %>% group_by(GEOID)

#try different states
pca.data <- scale(test[test$State == "NC", -c(1:2)])
pca.data <- pca.data[ , apply(pca.data, 2, function(x) !any(is.na(x)))]
test.pca <- princomp(pca.data)
summary(test.pca)
biplot(test.pca)


```

